---
title: "BoolNet Inference  (GSE41491)"
output:
  pdf_document: default
  html_notebook: default
  github_document: 
    df_print: paged
    html_preview: FALSE
    keep_html: TRUE
---

```{r error=TRUE, message=FALSE, warning=FALSE, include=FALSE, purl=FALSE, results='hide'}
## This chunk automatically generates a text .R version of this script when running within knitr.
input  = knitr::current_input()  # filename of input document
output = paste(tools::file_path_sans_ext(input), 'R', sep = '.')
try(knitr::purl(input,output,documentation=2,quiet=T), silent = TRUE)
# Avoid duplicate label error of knitr::purl
options(knitr.duplicate.label = 'allow')
# Code to browse the markdown file with renderized images.
knitr::opts_chunk$set(
  fig.path = "figs/GSE41491-"
)
```

Hypoxia transcriptomic time-series data in three different cancer cell lines (GSE41491)

Tumour hypoxia exhibits a highly dynamic spatial and temporal distribution and is associated with increased malignancy and poor prognosis.
Assessment of time-dependent gene-expression changes in response to hypoxia may thus provide additional biological insights and help with patient prognosis.

https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE41491

Exponentially growing prostate (DU145), colon (HT29) and breast (MCF7) carcinoma cells were seeded on glass dishes in McCoy, DMEM or RPMI media, respectively with 10% FCS.

All analyses were performed in R (v2.11.1). Data processing was performed for each cell line separately using RMA (affy package, v1.26.1) for pre-processing, and updated Entrez GeneID annotation (hgu133plus2hsentrezgcdf package, v13.0.0).

```{r message=FALSE, warning=FALSE}
packages_cran = c("igraph", "BoolNet", "BiocManager", "tidyverse", "fs")


# Install and load packages
package.check <- lapply(packages_cran, FUN = function(x) {
  if (!require(x, character.only = TRUE)) {
    install.packages(x, dependencies = TRUE)
    library(x, character.only = TRUE)
  }
})

# For oligo and ArrayExpress First install:
#install.packages('https://cran.r-project.org/src/contrib/Archive/ff/ff_2.2-14.tar.gz',repos=NULL)

if(!require("hgu133plus2hsentrezg.db")){ 
  install.packages('http://mbni.org/customcdf/13.0.0/entrezg.download/hgu133plus2hsentrezg.db_13.0.0.tar.gz',repos=NULL) 
  }

packages_bioconductor = c("Biobase", "GEOquery", "affyPLM", "ArrayExpress", "annotate")

# Install and load packages
package.check <- lapply(packages_bioconductor, FUN = function(x) {
  if (!require(x, character.only = TRUE)) {
    BiocManager::install(x, dependencies = TRUE)
    library(x, character.only = TRUE)
  }
})

rm(package.check, packages_bioconductor, packages_cran)

```


<!-- ```{r} -->

<!-- download_dir <- fs::path(".data_tmp") -->
<!-- if (!dir_exists(download_dir)) { dir_create(download_dir) } -->

<!-- GSE41491 <-getGEO("GSE41491", destdir = download_dir, GSEMatrix = T) -->

<!-- expr.GSE41491 <- as.matrix(exprs(normalize.ExpressionSet.quantiles(GSE41491[[1]]))) -->
<!-- data.GSE41491 <- pData(GSE41491[[1]]) -->

<!-- data.GSE41491 <- data.frame( -->
<!--                   codes = as.character(data.GSE41491$geo_accession), -->
<!--                   cell_line = gsub("[(]","",gsub("[)]","",substr(as.character(data.GSE41491$source_name_ch1),18,25))), -->
<!--                   time = substr(as.character(data.GSE41491$characteristics_ch1.1),18,19), -->
<!--                   condition = substr(as.character(data.GSE41491$characteristics_ch1),12,13)) -->

<!-- GPL14877 <- read_delim(".data_tmp/GPL14877.soft", -->
<!--     "\t", escape_double = FALSE, trim_ws = TRUE, -->
<!--     skip = 393) -->

<!-- ENTREZID <- as.character(unique(GPL14877$SPOT_ID)) -->
<!-- PROBES <- unique(rownames(expr.GSE41491)) -->

<!-- library(hgu133plus2hsentrezg.db) -->
<!-- SYMBOL <- getSYMBOL(PROBES, "hgu133plus2hsentrezg") -->
<!-- results <- as.data.frame(cbind(PROBES, SYMBOL)) -->

<!-- # Convert the probes to Symbol names -->
<!-- anno.GSE41491 <- AnnotationDbi::select(hgu133plus2.db, -->
<!--        keys = ENTREZID, -->
<!--        columns=c("SYMBOL", "GENENAME"), -->
<!--        keytype="ENTREZID") -->

<!-- anno.GSE41491 <- merge(anno.GSE41491, results, by="SYMBOL", all.x = T) %>% -->
<!--   filter(!is.na(SYMBOL)) %>% -->
<!--   mutate_all(as.character) %>% -->
<!--   rename_at(vars(c("SYMBOL", "ENTREZID", "GENENAME", "PROBES")), ~ c("symbol", "entrezid", "description", "probes")) -->

<!-- rm(download_dir, GSE41491, results, GPL14877, PROBES, SYMBOL, ENTREZID) -->

<!-- save.image("../data/data.GSE41491.Rdata") -->
<!-- ``` -->


# Load the pre-processed data

```{r}
load("../data/data.GSE41491.Rdata")
```


# Selecting the HIF Genes 
    
```{r}
# Selected genes from HIF Axis
hif.symbols <- c("TP53", "HIF1A", "EP300", "MDM2", "VBP1") 

hif.probes <- unique(anno.GSE41491$probes[anno.GSE41491$symbol %in% hif.symbols])

# Select the probes and genes
expr.GSE41491.hif <- data.frame(expr.GSE41491) %>% 
  rownames_to_column('probes') %>% 
  filter(probes %in% hif.probes) %>% 
  merge(anno.GSE41491[anno.GSE41491$symbol %in% hif.symbols, c("probes","symbol")], by = "probes") %>% 
  #distinct(symbol, .keep_all = TRUE) %>% # Take the first one
  dplyr::select(!(probes)) %>% 
  arrange(symbol)
  
```

```{r}
# Function to binarize according an consensus mean of probes, add the O2 state and rename columns 

binNet <- function(b){
  
  cols <- data.GSE41491$codes %in% names(b)
  
  binarizeTimeSeries(b[,-1], method="kmeans")$binarizedMeasurements  %>% 
  as.data.frame(.)  %>% 
  aggregate(., list(symbol = b$symbol), mean) %>%  # mean of binarized probes
  mutate_at(vars(-symbol), funs(ifelse(. >= 0.5, 1, 0))) %>%  # consensus with a bies to 1 (>= 0.5)
  rbind(., c("O2", 1,,0,0,0,0,0,0,0)) %>% 
    rename_at(vars(data.GSE41491$codes[cols] ),
            ~paste0(substr(data.GSE41491$condition[cols],1,2),".",
                    data.GSE41491$time[cols],".",
                    substr(data.GSE41491$cell_line[cols],1,2))) %>% 
  column_to_rownames("symbol")
  
}
```

Exponentially growing prostate (DU145), colon (HT29) and breast (MCF7) carcinoma cells were seeded on glass dishes in McCoy, DMEM or RPMI media, respectively with 10% FCS.

```{r}

breast1_MCF7 <- 
expr.GSE41491.hif %>% 
  dplyr::select(c("symbol", data.GSE41491$codes[data.GSE41491$cell_line == "MCF7"])) %>% arrange(symbol) %>% 
  rename_at(vars(data.GSE41491$codes[data.GSE41491$cell_line == "MCF7"]),
            ~paste0(data.GSE41491$condition[data.GSE41491$cell_line == "MCF7"],".",
                    data.GSE41491$time[data.GSE41491$cell_line == "MCF7"],".",
                    data.GSE41491$cell_line[data.GSE41491$cell_line == "MCF7"]))

knitr::kable(breast1_MCF7)

binarizeTimeSeries(breast1_MCF7[,-1], method="kmeans")$binarizedMeasurements  %>% 
  data.frame(.)  %>% 
  add_column(symbol = breast1_MCF7$symbol, .before=0) %>% 
  knitr::kable(.)
  
binarizeTimeSeries(breast1_MCF7[,-1], method="kmeans")$binarizedMeasurements  %>% 
  data.frame(.)  %>% 
  aggregate(., list(symbol = breast1_MCF7$symbol), mean) %>% 
  mutate_at(vars(-symbol), funs(ifelse(. >= 0.5, 1, 0))) %>% 
  rbind(., c("O2", 1,0,0,0,0,0,0,0)) %>% 
  knitr::kable(.)
  
```

# breast (MCF7) carcinoma cells 

```{r}

breast_MCF7 <- 
expr.GSE41491.hif %>% 
  dplyr::select(c("symbol", data.GSE41491$codes[data.GSE41491$cell_line == "MCF7"]))  %>% 
  binNet(.) 
knitr::kable(breast_MCF7)


# Individual net

net <- reconstructNetwork(breast_MCF7, method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)
plotNetworkWiring(net)
print(net)

```

# prostate (DU145) carcinoma cells 

```{r}

prostate_DU145 <- 
expr.GSE41491.hif %>% 
  dplyr::select(c("symbol", data.GSE41491$codes[data.GSE41491$cell_line == "DU145"]))  %>% 
  binNet(.) 
knitr::kable(prostate_DU145)


net <- reconstructNetwork(prostate_DU145, method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)
plotNetworkWiring(net)
print(net)

```

#  colon (HT29)  carcinoma cells 

```{r}

colon_MCF7 <- 
expr.GSE41491.hif %>% 
  dplyr::select(c("symbol", data.GSE41491$codes[data.GSE41491$cell_line == "HT29"]))  %>% 
  binNet(.) 
knitr::kable(colon_MCF7)


# Individual net

net <- reconstructNetwork(colon_MCF7, method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)
plotNetworkWiring(net)
print(net)

```


# All three cells prostate (DU145), colon (HT29) and breast (MCF7) carcinoma 

```{r}

net <- reconstructNetwork(list(breast_MCF7,prostate_DU145, colon_MCF7), method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)
plotNetworkWiring(net)
print(net)

```
