---
title: "BoolNet Inference  (E-GEOD-18494)"
output:
  pdf_document: default
  github_document: 
    df_print: paged
    html_preview: FALSE
    keep_html: TRUE
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_format = "all") })    
---
```{r error=TRUE, message=FALSE, warning=FALSE, include=FALSE, purl=FALSE, results='hide'}
## This chunk automatically generates a text .R version of this script when running within knitr.
input  = knitr::current_input()  # filename of input document
output = paste(tools::file_path_sans_ext(input), 'R', sep = '.')
try(knitr::purl(input,output,documentation=2,quiet=T), silent = TRUE)
# Avoid duplicate label error of knitr::purl
options(knitr.duplicate.label = 'allow')
# Code to browse the markdown file with renderized images.
knitr::opts_chunk$set(
  fig.path = "figs/EGEOD18494-"
)
```

Expression profiling of hypoxic HepG2 hepatoma, U87 glioma, and MDA-MB231 breast cancer cells: time course (E-GEOD-18494)

Analysis of expression changes of cultured HepG2 hepatoma, U87 glioma, and MDA-MB231 breast cancer cells subjected to hypoxia (0.5% O2) for 0, 4, 8, 12 hours . Results provide insight to cell type-specific response to hypoxia. HepG2 hepatoma, U87 glioma, and MDA-MB231 breast cancer cells were collected under normoxic conditions (~19% O2, 0 hours) and after 4, 8 and 12 hours of hypoxia treatment (0.5% O2). For each cell line, three replicates of total RNA at each time point were prepared using Trizol and submitted to the DFCI Microarray Core for labeling, hybridization to Affymetrix HG-U133Plus2 oligonucleotide arrays and image scanning.
    
https://www.ebi.ac.uk/arrayexpress/experiments/E-GEOD-18494/   


```{r message=FALSE, warning=FALSE}
packages_cran = c("igraph", "BoolNet", "BiocManager", "tidyverse", "fs", "ff", "RSQLite")


# Install and load packages
package.check <- lapply(packages_cran, FUN = function(x) {
  if (!require(x, character.only = TRUE)) {
    install.packages(x, dependencies = TRUE)
    library(x, character.only = TRUE)
  }
})

# For oligo First install:
#install.packages('https://cran.r-project.org/src/contrib/Archive/ff/ff_2.2-14.tar.gz',repos=NULL)

packages_bioconductor = c("Biobase", "GEOquery", "ArrayExpress", "oligo", "hgu133plus2.db")

# Install and load packages
package.check <- lapply(packages_bioconductor, FUN = function(x) {
  if (!require(x, character.only = TRUE)) {
    BiocManager::install(x, dependencies = TRUE)
    library(x, character.only = TRUE)
  }
})

rm(package.check, packages_bioconductor, packages_cran)

```

<!-- ```{r message=FALSE, warning=FALSE } -->
<!-- download_dir <- fs::path(".data_tmp") -->
<!-- if (!dir_exists(download_dir)) { -->
<!--     dir_create(download_dir) -->
<!--     EGEOD18494 <- ArrayExpress( "E-GEOD-18494", save=TRUE, path=download_dir) -->
<!-- } else { -->
<!--     EGEOD18494 <- ArrayExpress( "E-GEOD-18494", save=TRUE, path=download_dir) -->
<!-- } -->

<!-- data.EGEOD18494 <- Biobase::pData(EGEOD18494) -->

<!-- data.EGEOD18494 <- data.frame( -->
<!--                   codes = substr(data.EGEOD18494$Source.Name,1,9), -->
<!--                   cell_line = data.EGEOD18494$Characteristics..cell.line., -->
<!--                   time = data.EGEOD18494$Characteristics..time, -->
<!--                   condition = data.EGEOD18494$Characteristics..stress. -->
<!--                   ) -->
<!-- data.EGEOD18494 <- data.EGEOD18494[order(data.EGEOD18494$codes),] -->
<!-- data.EGEOD18494$rep <- rep(1:3, n= length(data.EGEOD18494$codes)) -->

<!-- # Normalisation -->
<!-- eset.EGEOD18494 <- oligo::rma(EGEOD18494,  normalize = TRUE) -->

<!-- expr.EGEOD18494 <- exprs(eset.EGEOD18494) -->

<!-- # Convert to a data.frame -->
<!-- expr.EGEOD18494 <- as.data.frame(as.ffdf(expr.EGEOD18494)) -->

<!-- colnames(expr.EGEOD18494) <- substr(colnames(expr.EGEOD18494),1,9) -->

<!-- # Convert the probes to Symbol names -->
<!-- anno.EGEOD18494 <- AnnotationDbi::select(hgu133plus2.db,  -->
<!--                                          keys=rownames(expr.EGEOD18494),  -->
<!--                                          columns=c("ENSEMBL", "SYMBOL", "GENENAME"),  -->
<!--                                          keytype="PROBEID") -->

<!-- colnames(anno.EGEOD18494) <- c("probes", "ensgene", "symbol", "description") -->

<!-- rm(download_dir, EGEOD18494, eset.EGEOD18494) -->

<!-- save.image("../data/data.EGEOD18494.Rdata") -->
<!-- ``` -->

# Load the pre-processed

```{r}
load("../data/data.EGEOD18494.Rdata")
```

# Selecting the HIF Genes 
    
```{r}

# Selecting genes from HIF Axis
hif.symbols <- c("TP53", "HIF1A", "EP300", "MDM2", "VHL")

hif.probes <- anno.EGEOD18494$probes[anno.EGEOD18494$symbol %in% hif.symbols]

# Select the probes and genes
expr.EGEOD18494.hif <- as.data.frame(expr.EGEOD18494) %>% 
  rownames_to_column('probes') %>% 
  filter(probes %in% hif.probes) %>% 
  merge(anno.EGEOD18494[anno.EGEOD18494$symbol %in% hif.symbols, c("probes","symbol")], by = "probes") %>% 
  #distinct(symbol, .keep_all = TRUE) %>% # Take the first one
  dplyr::select(!(probes)) 
  
```


```{r}
# Function to binarize according an consensus mean of probes, add the O2 state and rename columns 

binNet <- function(b){
  
  cols <- data.EGEOD18494$codes %in% names(b)
  
  binarizeTimeSeries(b[,-1], method="kmeans")$binarizedMeasurements  %>% 
  as.data.frame(.)  %>% 
  aggregate(., list(symbol = b$symbol), mean) %>%  # mean of binarized probes
  mutate_at(vars(-symbol), funs(ifelse(. >= 0.5, 1, 0))) %>%  # consensus with a bies to 1 (>= 0.5)
  rbind(., c("O2", 1,0,0,0)) %>% 
    rename_at(vars(data.EGEOD18494$codes[cols] ),
            ~paste0(substr(data.EGEOD18494$condition[cols],1,2),".",
                    data.EGEOD18494$time[cols],".",
                    substr(data.EGEOD18494$cell_line[cols],1,2), ".",
                    data.EGEOD18494$rep[cols])) %>% 
  column_to_rownames("symbol")
  
}
```

# Exemplifying the Binarization 

```{r}

cols <- (data.EGEOD18494$cell_line == "MDA-MB231 breast cancer" & data.EGEOD18494$rep == 1)

breast1x <- 
expr.EGEOD18494.hif %>% 
  dplyr::select(c("symbol", data.EGEOD18494$codes[cols])) %>% arrange(symbol) %>% 
  arrange(symbol) %>% 
  rename_at(vars(data.EGEOD18494$codes[cols]),
            ~paste0(substr(data.EGEOD18494$condition[cols],1,2),".",
                    data.EGEOD18494$time[cols],".",
                    substr(data.EGEOD18494$cell_line[cols],1,2)))

breast1x %>% 
  knitr::kable(.)

binarizeTimeSeries(breast1x[,-1], method="kmeans")$binarizedMeasurements  %>% 
  data.frame(.)  %>% 
  add_column(symbol = breast1x$symbol, .before=0) %>% 
  knitr::kable(.)
  
binarizeTimeSeries(breast1x[,-1], method="kmeans")$binarizedMeasurements  %>% 
  data.frame(.)  %>% 
  aggregate(., list(symbol = breast1x$symbol), mean) %>% 
  mutate_at(vars(-symbol), funs(ifelse(. >= 0.5, 1, 0))) %>% 
  rbind(., c("O2", 1,0,0,0)) %>% 
  knitr::kable(.)

```

# MDA-MB231 breast cancer

```{r}

cellline.rep1 <- (data.EGEOD18494$cell_line == "MDA-MB231 breast cancer" &  data.EGEOD18494$rep == 1)
cellline.rep2 <- (data.EGEOD18494$cell_line == "MDA-MB231 breast cancer" &  data.EGEOD18494$rep == 2)
cellline.rep3 <- (data.EGEOD18494$cell_line == "MDA-MB231 breast cancer" &  data.EGEOD18494$rep == 3)

breast1x <- 
expr.EGEOD18494.hif %>% 
  dplyr::select(c("symbol", data.EGEOD18494$codes[cellline.rep1])) %>% 
  binNet(.) 

breast1x %>% 
  knitr::kable(.)

breast2x <- 
expr.EGEOD18494.hif %>% 
  dplyr::select(c("symbol", data.EGEOD18494$codes[cellline.rep2])) %>% 
  binNet(.) 

breast2x  %>% 
  knitr::kable(.)

breast3x <- 
expr.EGEOD18494.hif %>% 
  dplyr::select(c("symbol", data.EGEOD18494$codes[cellline.rep3])) %>% 
  binNet(.) 

breast3x %>% 
  knitr::kable(.)

# All breast cancer nets merged:

net <- reconstructNetwork(list(breast1x, breast2x, breast3x), method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)
plotNetworkWiring(net)
print(net)

# Individual nets of each replica:

net <- reconstructNetwork(breast1x, method="bestfit", returnPBN=TRUE, readableFunctions=TRUE)
plotNetworkWiring(net)
print(net)

net <- reconstructNetwork(breast2x, method="bestfit", returnPBN=TRUE, readableFunctions=TRUE)
plotNetworkWiring(net)
print(net)

net <- reconstructNetwork(breast3x, method="bestfit", returnPBN=TRUE, readableFunctions=TRUE)
plotNetworkWiring(net)
print(net)
```


# HepG2 hepatoma

```{r}

cellline.rep1 <- (data.EGEOD18494$cell_line == "HepG2 hepatoma" &  data.EGEOD18494$rep == 1)
cellline.rep2 <- (data.EGEOD18494$cell_line == "HepG2 hepatoma" &  data.EGEOD18494$rep == 2)
cellline.rep3 <- (data.EGEOD18494$cell_line == "HepG2 hepatoma" &  data.EGEOD18494$rep == 3)

hepatoma1x <- 
expr.EGEOD18494.hif %>% 
  dplyr::select(c("symbol", data.EGEOD18494$codes[cellline.rep1]))  %>% 
  binNet(.) 

hepatoma1x %>% 
  knitr::kable(.)

hepatoma2x <- 
expr.EGEOD18494.hif %>% 
  dplyr::select(c("symbol", data.EGEOD18494$codes[cellline.rep2]))  %>% 
  binNet(.) 

hepatoma2x %>% 
  knitr::kable(.)

hepatoma3x <- 
expr.EGEOD18494.hif %>% 
  dplyr::select(c("symbol", data.EGEOD18494$codes[cellline.rep3]))  %>% 
  binNet(.) 

hepatoma3x %>% 
  knitr::kable(.)

# All nets hepatoma merged:

net <- reconstructNetwork(list(hepatoma1x, hepatoma2x, hepatoma3x), method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)
plotNetworkWiring(net)
print(net)

# Individual nets of each replica:

net <- reconstructNetwork(hepatoma1x, method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)
plotNetworkWiring(net)
print(net)

net <- reconstructNetwork(hepatoma2x, method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)
plotNetworkWiring(net)
print(net)

net <- reconstructNetwork(hepatoma3x, method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)
plotNetworkWiring(net)
print(net)
```

# U87 glioma

```{r}
cellline.rep1 <- (data.EGEOD18494$cell_line == "U87 glioma" &  data.EGEOD18494$rep == 1)
cellline.rep2 <- (data.EGEOD18494$cell_line == "U87 glioma" &  data.EGEOD18494$rep == 2)
cellline.rep3 <- (data.EGEOD18494$cell_line == "U87 glioma" &  data.EGEOD18494$rep == 3)

glioma1x <- 
expr.EGEOD18494.hif %>% 
  dplyr::select(c("symbol", data.EGEOD18494$codes[cellline.rep1]))  %>% 
  binNet(.) 

glioma1x %>% 
  knitr::kable(.)

glioma2x <- 
expr.EGEOD18494.hif %>% 
  dplyr::select(c("symbol", data.EGEOD18494$codes[cellline.rep1]))  %>% 
  binNet(.) 

glioma2x %>% 
  knitr::kable(.)

glioma3x <- 
expr.EGEOD18494.hif %>% 
  dplyr::select(c("symbol", data.EGEOD18494$codes[cellline.rep1]))  %>% 
  binNet(.) 

glioma3x %>% 
  knitr::kable(.)

# All glioma nets merged:

net <- reconstructNetwork(list(glioma1x, glioma2x, glioma3x), method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)
plotNetworkWiring(net)
print(net)

# Individual nets of each replica:

net <- reconstructNetwork(glioma1x, method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)
plotNetworkWiring(net)
print(net)

net <- reconstructNetwork(glioma2x, method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)
plotNetworkWiring(net)
print(net)

net <- reconstructNetwork(glioma3x, method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)
plotNetworkWiring(net)
print(net)
```

