---
title: "BoolNet Inference  (GSE41491)"
output:
  pdf_document: default
  github_document: 
    df_print: paged
    html_preview: FALSE
    keep_html: FALSE
---

```{r error=TRUE, message=FALSE, warning=FALSE, include=FALSE, purl=FALSE, results='hide'}
## This chunk automatically generates a text .R version of this script when running within knitr.
# input  = knitr::current_input()  # filename of input document
# output = paste(tools::file_path_sans_ext(input), 'R', sep = '.')
# try(knitr::purl(input,output,documentation=2,quiet=T), silent = TRUE)
# Avoid duplicate label error of knitr::purl
options(knitr.duplicate.label = 'allow')
# Code to browse the markdown file with renderized images.
knitr::opts_chunk$set(
  fig.path = "figs/GSE41491-"
)
```

Hypoxia transcriptomic time-series data in three different cancer cell lines (GSE41491)

Tumour hypoxia exhibits a highly dynamic spatial and temporal distribution and is associated with increased malignancy and poor prognosis. Exponentially growing prostate (DU145), colon (HT29) and breast (MCF7) carcinoma cells were seeded on glass dishes in McCoy, DMEM or RPMI media, respectively with 10% FCS.

https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE41491




```{r message=FALSE, warning=FALSE}
packages_cran = c("igraph", "BoolNet", "BiocManager", "tidyverse", "fs")


# Install and load packages
package.check <- lapply(packages_cran, FUN = function(x) {
  if (!require(x, character.only = TRUE)) {
    install.packages(x, dependencies = TRUE)
    library(x, character.only = TRUE)
  }
})

if(!require("hgu133plus2hsentrezg.db")){ 
  install.packages('http://mbni.org/customcdf/13.0.0/entrezg.download/hgu133plus2hsentrezg.db_13.0.0.tar.gz',repos=NULL) 
  }

packages_bioconductor = c("Biobase", "GEOquery", "affyPLM", "ArrayExpress", "annotate")

# Install and load packages
package.check <- lapply(packages_bioconductor, FUN = function(x) {
  if (!require(x, character.only = TRUE)) {
    BiocManager::install(x, dependencies = TRUE)
    library(x, character.only = TRUE)
  }
})

rm(package.check, packages_bioconductor, packages_cran)

```


<!-- ```{r} -->

<!-- download_dir <- fs::path(".data_tmp") -->
<!-- if (!dir_exists(download_dir)) { dir_create(download_dir) } -->

<!-- GSE41491 <-getGEO("GSE41491", destdir = download_dir, GSEMatrix = T) -->

<!-- expr.GSE41491 <- as.matrix(exprs(normalize.ExpressionSet.quantiles(GSE41491[[1]]))) -->
<!-- data.GSE41491 <- pData(GSE41491[[1]]) -->

<!-- data.GSE41491 <- data.frame( -->
<!--                   codes = as.character(data.GSE41491$geo_accession), -->
<!--                   cell_line = gsub("[(]","",gsub("[)]","",substr(as.character(data.GSE41491$source_name_ch1),18,25))), -->
<!--                   time = substr(as.character(data.GSE41491$characteristics_ch1.1),18,19), -->
<!--                   condition = substr(as.character(data.GSE41491$characteristics_ch1),12,13)) -->

<!-- GPL14877 <- read_delim(".data_tmp/GPL14877.soft", -->
<!--     "\t", escape_double = FALSE, trim_ws = TRUE, -->
<!--     skip = 393) -->

<!-- ENTREZID <- as.character(unique(GPL14877$SPOT_ID)) -->
<!-- PROBES <- unique(rownames(expr.GSE41491)) -->

<!-- library(hgu133plus2hsentrezg.db) -->
<!-- SYMBOL <- getSYMBOL(PROBES, "hgu133plus2hsentrezg") -->
<!-- results <- as.data.frame(cbind(PROBES, SYMBOL)) -->

<!-- # Convert the probes to Symbol names -->
<!-- anno.GSE41491 <- AnnotationDbi::select(hgu133plus2.db, -->
<!--        keys = ENTREZID, -->
<!--        columns=c("SYMBOL", "GENENAME"), -->
<!--        keytype="ENTREZID") -->

<!-- anno.GSE41491 <- merge(anno.GSE41491, results, by="SYMBOL", all.x = T) %>% -->
<!--   filter(!is.na(SYMBOL)) %>% -->
<!--   mutate_all(as.character) %>% -->
<!--   rename_at(vars(c("SYMBOL", "ENTREZID", "GENENAME", "PROBES")), ~ c("symbol", "entrezid", "description", "probes")) -->

<!-- rm(download_dir, GSE41491, results, GPL14877, PROBES, SYMBOL, ENTREZID) -->

<!-- save.image("../data/data.GSE41491.Rdata") -->
<!-- ``` -->

# Load the pre-processed data

```{r}
load("../data/data.GSE41491.Rdata")
cols <- colnames(expr.GSE41491)
rows <- rownames(expr.GSE41491)
expr.GSE41491 <- data.frame(matrix(effectsize::normalize(as.matrix(expr.GSE41491)), ncol = length(cols), nrow = length(rows) ))
colnames(expr.GSE41491) <- cols
rownames(expr.GSE41491) <- rows 
```

# Selecting the HIF Genes 
    
```{r}
# Selected genes from HIF Axis
hif.symbols <- c("TP53", "HIF1A", "EP300", "MDM2", "VBP1") 
hif.probes <- unique(anno.GSE41491$probes[anno.GSE41491$symbol %in% hif.symbols])

# Select the probes and genes
expr.GSE41491.hif <- data.frame(expr.GSE41491) %>% 
  rownames_to_column('probes') %>% 
  filter(probes %in% hif.probes) %>% 
  merge(anno.GSE41491[anno.GSE41491$symbol %in% hif.symbols, c("probes","symbol")], by = "probes") %>% 
  dplyr::select(!(probes)) %>%  arrange(symbol)

expr.GSE41491.hif$symbol[expr.GSE41491.hif$symbol == "VBP1"] <- "VHL"
```

```{r  include=FALSE}
# Function to binarize according an consensus mean of probes, add the O2 state and rename columns 
binNet <- function(b){
  
  cols <- data.GSE41491$codes %in% names(b)
  
  binarizeTimeSeries(b[,-1], method="kmeans")$binarizedMeasurements  %>% 
  as.data.frame(.)  %>% 
  aggregate(., list(symbol = b$symbol), mean) %>%  # mean of binarized probes
  mutate_at(vars(-symbol), funs(ifelse(. >= 0.5, 1, 0))) %>%  # consensus with a bies to 1 (>= 0.5)
  rbind(., c("O2", 1,0,0,0,0,0,0,0)) %>% 
    rename_at(vars(data.GSE41491$codes[cols] ),
            ~paste0(substr(data.GSE41491$condition[cols],1,2),".",
                    data.GSE41491$time[cols],".",
                    substr(data.GSE41491$cell_line[cols],1,2))) %>% 
  column_to_rownames("symbol")
}

breast_MCF7 <- 
expr.GSE41491.hif %>% 
  dplyr::select(c("symbol", data.GSE41491$codes[data.GSE41491$cell_line == "MCF7"]))  %>% 
  binNet(.) 
knitr::kable(breast_MCF7)

prostate_DU145 <- 
expr.GSE41491.hif %>% 
  dplyr::select(c("symbol", data.GSE41491$codes[data.GSE41491$cell_line == "DU145"]))  %>% 
  binNet(.) 
knitr::kable(prostate_DU145)

colon_HT29 <- 
expr.GSE41491.hif %>% 
  dplyr::select(c("symbol", data.GSE41491$codes[data.GSE41491$cell_line == "HT29"]))  %>% 
  binNet(.) 
knitr::kable(colon_HT29)
```

# breast (MCF7) carcinoma cells - 8 time-points

```{r}

breast1_MCF7 <- 
expr.GSE41491.hif %>% 
  dplyr::select(c("symbol", data.GSE41491$codes[data.GSE41491$cell_line == "MCF7"])) %>% arrange(symbol) %>% 
  rename_at(vars(data.GSE41491$codes[data.GSE41491$cell_line == "MCF7"]),
            ~paste0(data.GSE41491$condition[data.GSE41491$cell_line == "MCF7"],".",
                    data.GSE41491$time[data.GSE41491$cell_line == "MCF7"],".",
                    data.GSE41491$cell_line[data.GSE41491$cell_line == "MCF7"]))

knitr::kable(breast1_MCF7)

binarizeTimeSeries(breast1_MCF7[,-1], method="kmeans")$binarizedMeasurements  %>% 
  data.frame(.)  %>% 
  add_column(symbol = breast1_MCF7$symbol, .before=0) %>% 
  knitr::kable(.)
  
binarizeTimeSeries(breast1_MCF7[,-1], method="kmeans")$binarizedMeasurements  %>% 
  data.frame(.)  %>% 
  aggregate(., list(symbol = breast1_MCF7$symbol), mean) %>% 
  mutate_at(vars(-symbol), funs(ifelse(. >= 0.5, 1, 0))) %>% 
  rbind(., c("O2", 1,0,0,0,0,0,0,0)) %>% 
  knitr::kable(.)
  
```

# prostate (DU145) carcinoma cells - 8 time-points

```{r}
prostate_DU145 %>% 
  knitr::kable(.)
```

# colon (HT29) carcinoma cells - 8 time-points

```{r}
colon_HT29 %>% 
  knitr::kable(.)
```

```{r}
# breast (MCF7) carcinoma cells - 8 time-points
breast_MCF7.net <- reconstructNetwork(breast_MCF7, method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)
breast_MCF7.p <- plotNetworkWiring(breast_MCF7.net, plotIt=F)

# prostate (DU145) carcinoma cells - 8 time-points
prostate_DU145.net <- reconstructNetwork(prostate_DU145, method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)
prostate_DU145.p <- plotNetworkWiring(prostate_DU145.net, plotIt=F)

# colon (HT29) carcinoma cells - 8 time-points
colon_HT29.net <- reconstructNetwork(colon_HT29, method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)
colon_HT29.p <- plotNetworkWiring(colon_HT29.net, plotIt=F)
```


```{r}
par(mfrow = c(1,3))
plot(breast_MCF7.p, vertex.label.color="#440154ff", vertex.color="lightblue", vertex.frame.color="white", layout=layout_in_circle, edge.curved=.3,
     main="breast (MCF7)\n 8 steps")
plot(prostate_DU145.p, vertex.label.color="#440154ff", vertex.color="lightblue", vertex.frame.color="white", layout=layout_in_circle, edge.curved=.3,
     main="prostate (DU145)\n 8 steps")
plot(colon_HT29.p, vertex.label.color="#440154ff", vertex.color="lightblue", vertex.frame.color="white", layout=layout_in_circle, edge.curved=.3,
     main="colon (HT29)\n 8 steps")
```


<!-- ```{r} -->
<!-- all.nets <- reconstructNetwork(list(breast_MCF7, prostate_DU145, colon_HT29), method="bestfit",returnPBN=TRUE,readableFunctions=TRUE) -->
<!-- plotNetworkWiring(net.all, vertex.label.color="#440154ff", vertex.color="lightblue", vertex.frame.color="white", layout=layout_in_circle, edge.curved=.3, -->
<!--      main="all lines \n breast, prostate, colon - 4 steps") -->
<!-- ``` -->


```{r}
# breast (MCF7) carcinoma cells - 8 time-points
print(breast_MCF7.net)

# prostate (DU145) carcinoma cells - 8 time-points
print(prostate_DU145.net)

# colon (HT29) carcinoma cells - 8 time-points
print(colon_HT29.net)

# all lines - breast, prostate, colon - 4 steps
#print(all.nets)
```

