---
title: "BoolNet Inference  (GSE41491)"
output:
  pdf_document: default
  html_notebook: default
  github_document: 
    df_print: paged
    html_preview: FALSE
    keep_html: TRUE
---

```{r error=TRUE, message=FALSE, warning=FALSE, include=FALSE, purl=FALSE, results='hide'}
## This chunk automatically generates a text .R version of this script when running within knitr.
input  = knitr::current_input()  # filename of input document
output = paste(tools::file_path_sans_ext(input), 'R', sep = '.')
try(knitr::purl(input,output,documentation=2,quiet=T), silent = TRUE)
# Avoid duplicate label error of knitr::purl
options(knitr.duplicate.label = 'allow')
# Code to browse the markdown file with renderized images.
knitr::opts_chunk$set(
  fig.path = "figs/GSE41491-"
)
```

Hypoxia transcriptomic time-series data in three different cancer cell lines (GSE41491)

Tumour hypoxia exhibits a highly dynamic spatial and temporal distribution and is associated with increased malignancy and poor prognosis.
Assessment of time-dependent gene-expression changes in response to hypoxia may thus provide additional biological insights and help with patient prognosis.

https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE41491

Exponentially growing prostate (DU145), colon (HT29) and breast (MCF7) carcinoma cells were seeded on glass dishes in McCoy, DMEM or RPMI media, respectively with 10% FCS.

All analyses were performed in R (v2.11.1). Data processing was performed for each cell line separately using RMA (affy package, v1.26.1) for pre-processing, and updated Entrez GeneID annotation (hgu133plus2hsentrezgcdf package, v13.0.0).

```{r message=FALSE, warning=FALSE}
packages_cran = c("igraph", "BoolNet", "BiocManager", "tidyverse", "fs", "ff")


# Install and load packages
package.check <- lapply(packages_cran, FUN = function(x) {
  if (!require(x, character.only = TRUE)) {
    install.packages(x, dependencies = TRUE)
    library(x, character.only = TRUE)
  }
})

# For oligo and ArrayExpress First install:
#install.packages('https://cran.r-project.org/src/contrib/Archive/ff/ff_2.2-14.tar.gz',repos=NULL)

packages_bioconductor = c("Biobase", "GEOquery", "oligo", "ArrayExpress", "preprocessCore")

# Install and load packages
package.check <- lapply(packages_bioconductor, FUN = function(x) {
  if (!require(x, character.only = TRUE)) {
    BiocManager::install(x, dependencies = TRUE)
    library(x, character.only = TRUE)
  }
})

rm(package.check, packages_bioconductor, packages_cran)

```

```{r}

download_dir <- fs::path(".data_tmp")
if (!dir_exists(download_dir)) { dir_create(download_dir) }  

GSE41491 <-getGEO("GSE41491", destdir = download_dir, GSEMatrix = T)

expr.GSE41491 <- exprs(GSE41491[[1]])
prob.GSE41491 <- unique(rownames(expr.GSE41491))
data.GSE41491 <- pData(GSE41491[[1]])

data.GSE41491 <- data.frame(
                  codes = as.character(data.GSE41491$geo_accession),
                  cell_line = substr(as.character(data.GSE41491$source_name_ch1),18,25),
                  time = substr(as.character(data.GSE41491$characteristics_ch1.1),18,19),
                  condition = substr(as.character(data.GSE41491$characteristics_ch1),12,13))

# Convert the probes to Symbol names
#Affymetrix Human Genome U133 Plus 2.0 Array [Brainarray Version 15.0.0, HGU133Plus2_Hs_ENTREZG]
#  load/install the package
if(!require("hgu133plus2.db")) BiocManager::install("hgu133plus2.db")

# The below function call will return a datafram with probe_id, gene symbol
# and ŕefgene_id for your data

library(readr)
GPL14877 <- read_delim("workflow_boolnet/.data_tmp/GPL14877.soft", 
    "\t", escape_double = FALSE, trim_ws = TRUE, 
    skip = 393)

anno.GSE41491 <- AnnotationDbi::select(hgu133plus2.db, 
       keys = prob.GSE41491, 
       columns=c("ENTREZID", "ENSEMBL", "SYMBOL", "GENENAME"), 
       keytype="PROBEID")

colnames(anno.GSE41491) <- c("probes", "ensgene", "symbol", "description")

rm(download_dir, GSE41491,  prob.GSE41491)

```


# Selecting the HIF Genes 
    
```{r}
# Genes from Boolean Network:    
# HIF1a, HIF2a, p53, BNIP3, VEGF, cMyc, Oct4, cdc20, cycA, cycB, cycE, cycD, p27, Rb, E2F, cdh1, mdm2, BAD, BclX

# hif.symbols <- c("HIF1A", "HIF1", "PASD8", "MOP1", "EPAS1", "HIF2A", "HLF", "PASD2", "MOP2", "VEGFA", "VPF", "MVCD1", "VEGF-A", "TP53", "P53", "MYC", "C-Myc", "POU5F1", "OCT3", "OTF3", "CDC20", "P55CDC", "CDC20A", "CCNA1", "CCNA1", "CCNA2", "CCN1", "CCNA", "CCNB1", "CCNB", "CCNB1", "CCNB2", "CCNB2", "HsT17299", "CCND1", "PRAD1", "CCND2", "MPPH3", "CCNE1", "CCNE1", "CCNE", "PCCNE1", "CCNE2", "CCNE2", "CCNE", "PCCNE1", "CDKN1B", "KIP1", "P27KIP1", "CDKN4MEN4", "RB1", "PPP1R130", "Pp110", "E2F1", "RBAP-1", "BNIP3", "NIP3", "BCL2", "MCL1", "BCL2L3", "CDH1", "CD324", "UBE2C", "BAD", "BCL2L8", "VHL", "PVHL", "VHL1", "MDM2", "HDM2", "EP300", "P300", "KAT3B")

# Selected genes from HIF Axis
hif.symbols <- c("TP53", "HIF1A", "EP300", "MDM2", "VHL")

hif.probes <- anno.GSE41491$probes[anno.GSE41491$symbol %in% hif.symbols]

# Select the probes and genes
expr.GSE41491.hif <- data.frame(expr.GSE41491) %>% 
  rownames_to_column('probes') %>% 
  filter(probes %in% hif.probes) %>% 
  merge(anno.GSE41491[anno.GSE41491$symbol %in% hif.symbols, c("probes","symbol")], by = "probes") %>% 
  #distinct(symbol, .keep_all = TRUE) %>% # Take the first one
  dplyr::select(!(probes)) %>% 
  arrange(symbol)
  

# Function to binarize according an consensus mean of probes, add the O2 state and rename columns 
binNet <- function(b){
  binarizeTimeSeries(b[,-5], method="kmeans")$binarizedMeasurements  %>% 
  data.frame(.)  %>% 
  aggregate(., list(symbol = b$symbol), mean) %>% 
  mutate_at(vars(-symbol), funs(ifelse(. > 0.4, 1, 0))) %>% 
  rbind(., c("O2", 1,0,0,0)) %>% 
    rename_at(vars(data.GSE41491$codes[data.GSE41491$codes %in% names(b)]),
            ~paste0(data.GSE41491$condition[data.GSE41491$codes %in% names(b)],".",
                    data.GSE41491$time[data.GSE41491$codes %in% names(b)],".",
                    data.GSE41491$rep[data.GSE41491$codes %in% names(b)])) %>% 
  column_to_rownames("symbol")
}


```

Exponentially growing prostate (DU145), colon (HT29) and breast (MCF7) carcinoma cells were seeded on glass dishes in McCoy, DMEM or RPMI media, respectively with 10% FCS.

```{r}

breast1_MCF7 <- 
expr.GSE41491.hif %>% 
  dplyr::select(c(data.GSE41491$codes[data.GSE41491$rep == 1], "symbol")) %>% arrange(symbol)

names(breast1_MCF7) <- c("Norm..1","Hypo.16h.1","Hypo.32h.1","Hypo.48h.1", "symbol")

knitr::kable(breast1_MCF7[, c("symbol", "Norm..1","Hypo.16h.1","Hypo.32h.1","Hypo.48h.1")])

binarizeTimeSeries(breast1_MCF7[,-5], method="kmeans")$binarizedMeasurements  %>% 
  data.frame(.)  %>% 
  add_column(symbol = breast1_MCF7$symbol) %>%   dplyr::select(c("symbol", "Norm..1","Hypo.16h.1","Hypo.32h.1","Hypo.48h.1"))  %>% 
  knitr::kable(.)
  
binarizeTimeSeries(breast1_MCF7[,-5], method="kmeans")$binarizedMeasurements  %>% 
  data.frame(.)  %>% 
  aggregate(., list(symbol = breast1_MCF7$symbol), mean) %>% 
  mutate_at(vars(-symbol), funs(ifelse(. > 0.4, 1, 0))) %>% 
  rbind(., c("O2", 1,0,0,0)) %>% 
  knitr::kable(.)
  
```

# MDA-MB231 breast cancer

```{r}

breast1_MCF7 <- 
expr.GSE41491.hif %>% 
  dplyr::select(c(data.GSE41491$codes[data.GSE41491$rep == 1], "symbol"))  %>% 
  binNet(.) 
knitr::kable(breast1_MCF7)


breast2_MCF7 <- 
expr.GSE41491.hif %>% 
  dplyr::select(c(data.GSE41491$codes[data.GSE41491$rep == 2], "symbol"))  %>% 
  binNet(.) 
knitr::kable(breast2_MCF7)


breast3_MCF7 <- 
expr.GSE41491.hif %>% 
  dplyr::select(c(data.GSE41491$codes[data.GSE41491$rep == 3], "symbol"))  %>% 
  binNet(.) 
knitr::kable(breast3_MCF7)

# All breast cancer nets merged:

net <- reconstructNetwork(list(breast1_MCF7, breast2_MCF7, breast3_MCF7), method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)
plotNetworkWiring(net)
print(net)

# Individual nets of each replica:

net <- reconstructNetwork(breast1_MCF7, method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)
plotNetworkWiring(net)
print(net)

net <- reconstructNetwork(breast2_MCF7, method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)
plotNetworkWiring(net)
print(net)

net <- reconstructNetwork(breast3_MCF7, method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)
plotNetworkWiring(net)
print(net)
```


