---
title: "BoolNet Inference HepG2 hepatoma, U87 glioma, and MDA-MB231 breast cancer (E-GEOD-18494)"
output:
  pdf_document: default
  github_document: 
    df_print: paged
    html_preview: FALSE
    keep_html: FALSE
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_format = "all") })    
---
```{r error=TRUE, message=FALSE, warning=FALSE, include=FALSE, purl=FALSE, results='hide'}
## This chunk automatically generates a text .R version of this script when running within knitr.
input  = knitr::current_input()  # filename of input document
output = paste(tools::file_path_sans_ext(input), 'R', sep = '.')
try(knitr::purl(input,output,documentation=2,quiet=T), silent = TRUE)
# Avoid duplicate label error of knitr::purl
options(knitr.duplicate.label = 'allow')
# Code to browse the markdown file with renderized images.
knitr::opts_chunk$set(
  fig.path = "figs/EGEOD18494-"
)
```

Expression profiling of hypoxic HepG2 hepatoma, U87 glioma, and MDA-MB231 breast cancer cells: time course (E-GEOD-18494)

Analysis of expression changes of cultured HepG2 hepatoma, U87 glioma, and MDA-MB231 breast cancer cells subjected to hypoxia (0.5% O2) for 0, 4, 8, 12 hours . Results provide insight to cell type-specific response to hypoxia. HepG2 hepatoma, U87 glioma, and MDA-MB231 breast cancer cells were collected under normoxic conditions (~19% O2, 0 hours) and after 4, 8 and 12 hours of hypoxia treatment (0.5% O2). For each cell line, three replicates of total RNA at each time point were prepared using Trizol and submitted to the DFCI Microarray Core for labeling, hybridization to Affymetrix HG-U133Plus2 oligonucleotide arrays and image scanning.
    
https://www.ebi.ac.uk/arrayexpress/experiments/E-GEOD-18494/   


```{r message=FALSE, warning=FALSE}
packages_cran = c("igraph", "BoolNet", "BiocManager", "tidyverse", "fs")
# Install and load packages
package.check <- lapply(packages_cran, FUN = function(x) {
  if (!require(x, character.only = TRUE)) {
    install.packages(x, dependencies = TRUE)
    library(x, character.only = TRUE)
  }
})
packages_bioconductor = c("Biobase", "GEOquery", "vsn", "hgu133plus2.db")
# Install and load packages
package.check <- lapply(packages_bioconductor, FUN = function(x) {
  if (!require(x, character.only = TRUE)) {
    BiocManager::install(x, dependencies = TRUE)
    library(x, character.only = TRUE)
  }
})

rm(package.check, packages_bioconductor, packages_cran)

```

<!-- ```{r message=FALSE, warning=FALSE } -->
<!-- download_dir <- fs::path(".data_tmp") -->
<!-- if (!dir_exists(download_dir)) { -->
<!--     dir_create(download_dir) -->
<!--     EGEOD18494 <- ArrayExpress( "E-GEOD-18494", save=TRUE, path=download_dir) -->
<!-- } else { -->
<!--     EGEOD18494 <- ArrayExpress( "E-GEOD-18494", save=TRUE, path=download_dir) -->
<!-- } -->

<!-- data.EGEOD18494 <- Biobase::pData(EGEOD18494) -->

<!-- data.EGEOD18494 <- data.frame( -->
<!--                   codes = substr(data.EGEOD18494$Source.Name,1,9), -->
<!--                   cell_line = data.EGEOD18494$Characteristics..cell.line., -->
<!--                   time = data.EGEOD18494$Characteristics..time, -->
<!--                   condition = data.EGEOD18494$Characteristics..stress. -->
<!--                   ) -->
<!-- data.EGEOD18494 <- data.EGEOD18494[order(data.EGEOD18494$codes),] -->
<!-- data.EGEOD18494$rep <- rep(1:3, n= length(data.EGEOD18494$codes)) -->

<!-- # Normalisation -->
<!-- eset.EGEOD18494 <- oligo::rma(EGEOD18494,  normalize = TRUE) -->

<!-- expr.EGEOD18494 <- exprs(eset.EGEOD18494) -->

<!-- # Convert to a data.frame -->
<!-- expr.EGEOD18494 <- as.data.frame(as.ffdf(expr.EGEOD18494)) -->

<!-- colnames(expr.EGEOD18494) <- substr(colnames(expr.EGEOD18494),1,9) -->

<!-- # Convert the probes to Symbol names -->
<!-- anno.EGEOD18494 <- AnnotationDbi::select(hgu133plus2.db,  -->
<!--                                          keys=rownames(expr.EGEOD18494),  -->
<!--                                          columns=c("ENSEMBL", "SYMBOL", "GENENAME"),  -->
<!--                                          keytype="PROBEID") -->

<!-- colnames(anno.EGEOD18494) <- c("probes", "ensgene", "symbol", "description") -->

<!-- rm(download_dir, EGEOD18494, eset.EGEOD18494) -->

<!-- save.image("../data/data.EGEOD18494.Rdata") -->
<!-- ``` -->

# Load the pre-processed

```{r}
load("../data/data.EGEOD18494.Rdata")
eset <- ExpressionSet(assayData = as.matrix(expr.EGEOD18494), 
                      probeNames = row.names(expr.EGEOD18494))
expr.EGEOD18494 <- exprs(justvsn(eset))
```

# Selecting the HIF Genes 
    
```{r}

# Selecting genes from HIF Axis
hif.symbols <- c("TP53", "HIF1A", "EP300", "MDM2", "VHL")

hif.probes <- anno.EGEOD18494$probes[anno.EGEOD18494$symbol %in% hif.symbols]

# Select the probes and genes
expr.EGEOD18494.hif <- as.data.frame(expr.EGEOD18494) %>% 
  rownames_to_column('probes') %>% 
  filter(probes %in% hif.probes) %>% 
  merge(anno.EGEOD18494[anno.EGEOD18494$symbol %in% hif.symbols, c("probes","symbol")], by = "probes") %>% 
  #distinct(symbol, .keep_all = TRUE) %>% # Take the first one
  dplyr::select(!(probes)) 
  
```


```{r}
# Function to binarize according an consensus mean of probes, add the O2 state and rename columns

binNet <- function(b){

  cols <- data.EGEOD18494$codes %in% names(b)

  binarizeTimeSeries(b[,-1], method="kmeans")$binarizedMeasurements  %>%
  as.data.frame(.)  %>%
  aggregate(., list(symbol = b$symbol), mean) %>%  # mean of binarized probes
  mutate_at(vars(-symbol), funs(ifelse(. >= 0.5, 1, 0))) %>%  # consensus with a bies to 1 (>= 0.5)
  #rbind(., c("O2", 1,0,0,0)) %>%
    rename_at(vars(data.EGEOD18494$codes[cols] ),
            ~paste0(substr(data.EGEOD18494$condition[cols],1,2),".",
                    data.EGEOD18494$time[cols],".",
                    substr(data.EGEOD18494$cell_line[cols],1,2), ".",
                    data.EGEOD18494$rep[cols])) %>%
  column_to_rownames("symbol")

}

```


```{r}
# Function to calculate the mean and binarize after, according an consensus mean of probes 

meanBinNet <- function(b){
  
  cols <- data.EGEOD18494$codes %in% names(b)
  
  b <-b %>% 
  rename_at(vars(data.EGEOD18494$codes[cols] ),
            ~paste0(substr(data.EGEOD18494$condition[cols],1,2),".",
                    data.EGEOD18494$time[cols],".",
                    substr(data.EGEOD18494$cell_line[cols],1,2), ".",
                    data.EGEOD18494$rep[cols])) %>% 
  mutate(no.ctrl = rowMeans(dplyr::select(.,starts_with("no.control")), na.rm = TRUE)) %>% 
  mutate(hy.4h = rowMeans(dplyr::select(.,starts_with("hy.4h")), na.rm = TRUE)) %>% 
  mutate(hy.8h = rowMeans(dplyr::select(.,starts_with("hy.8h")), na.rm = TRUE)) %>% 
  mutate(hy.12h = rowMeans(dplyr::select(.,starts_with("hy.12h")), na.rm = TRUE)) %>%
  dplyr::select(c("symbol", "no.ctrl", "hy.4h", "hy.8h", "hy.12h")) 
  
  binarizeTimeSeries(b[,-1], method="kmeans")$binarizedMeasurements  %>% 
  as.data.frame(.)  %>% 
  aggregate(., list(symbol = b$symbol), mean) %>%  # mean of binarized probes
  mutate_at(vars(-symbol), funs(ifelse(. >= 0.5, 1, 0))) %>%  # consensus with a bies to 1 (>= 0.5)
  #rbind(., c("O2", 1,0,0,0)) %>% 
  column_to_rownames("symbol")
  
}

```

# Exemplifying the Binarization 

```{r}

cols <- (data.EGEOD18494$cell_line == "MDA-MB231 breast cancer" & data.EGEOD18494$rep == 1)

breast1x <- 
expr.EGEOD18494.hif %>% 
  dplyr::select(c("symbol", data.EGEOD18494$codes[cols])) %>% arrange(symbol) %>% 
  arrange(symbol) %>% 
  rename_at(vars(data.EGEOD18494$codes[cols]),
            ~paste0(substr(data.EGEOD18494$condition[cols],1,2),".",
                    data.EGEOD18494$time[cols],".",
                    substr(data.EGEOD18494$cell_line[cols],1,2)))

breast1x %>% 
  knitr::kable(.)

binarizeTimeSeries(breast1x[,-1], method="kmeans")$binarizedMeasurements  %>% 
  data.frame(.)  %>% 
  add_column(symbol = breast1x$symbol, .before=0) %>% 
  knitr::kable(.)
  
binarizeTimeSeries(breast1x[,-1], method="kmeans")$binarizedMeasurements  %>% 
  data.frame(.)  %>% 
  aggregate(., list(symbol = breast1x$symbol), mean) %>% 
  mutate_at(vars(-symbol), funs(ifelse(. >= 0.5, 1, 0))) %>% 
  #rbind(., c("O2", 1,0,0,0)) %>% 
  knitr::kable(.)


```

# MDA-MB231 breast cancer


```{r}

cellline.breast <- (data.EGEOD18494$cell_line == "MDA-MB231 breast cancer")
                  
breast.meanBin <- 
expr.EGEOD18494.hif %>% 
  dplyr::select(c("symbol", data.EGEOD18494$codes[cellline.breast])) %>% meanBinNet(.) 

breast.meanBin %>% 
  knitr::kable(.)
```

```{r}
cellline.rep1 <- (data.EGEOD18494$cell_line == "MDA-MB231 breast cancer" &  data.EGEOD18494$rep == 1)
cellline.rep2 <- (data.EGEOD18494$cell_line == "MDA-MB231 breast cancer" &  data.EGEOD18494$rep == 2)
cellline.rep3 <- (data.EGEOD18494$cell_line == "MDA-MB231 breast cancer" &  data.EGEOD18494$rep == 3)

breast1x <-
expr.EGEOD18494.hif %>%
  dplyr::select(c("symbol", data.EGEOD18494$codes[cellline.rep1])) %>% binNet(.)

# breast1x %>% knitr::kable(.)

breast2x <-
expr.EGEOD18494.hif %>%
  dplyr::select(c("symbol", data.EGEOD18494$codes[cellline.rep2])) %>% binNet(.)

# breast2x  %>% knitr::kable(.)

breast3x <-
expr.EGEOD18494.hif %>%
  dplyr::select(c("symbol", data.EGEOD18494$codes[cellline.rep3])) %>% binNet(.)

# breast3x %>% knitr::kable(.)

breast.mean <-
cbind(breast1x,breast2x,breast3x) %>%
  tibble::rownames_to_column('gene') %>%
  mutate_at(vars(-gene), as.numeric) %>%
  mutate(no.ctrl = rowMeans(dplyr::select(.,starts_with("no.control")), na.rm = TRUE)) %>%
  mutate(hy.4h = rowMeans(dplyr::select(.,starts_with("hy.4h")), na.rm = TRUE)) %>%
  mutate(hy.8h = rowMeans(dplyr::select(.,starts_with("hy.8h")), na.rm = TRUE)) %>%
  mutate(hy.12h = rowMeans(dplyr::select(.,starts_with("hy.12h")), na.rm = TRUE)) %>%
  dplyr::select(c("no.ctrl", "hy.4h", "hy.8h", "hy.12h", "gene")) %>%
  mutate_at(c("no.ctrl", "hy.4h", "hy.8h", "hy.12h"), funs(ifelse(. >= 0.5, 1, 0)))  %>%  # consensus with a bies to 1 (>= 0.5)
  tibble::column_to_rownames('gene')

# breast.mean %>%  
#    knitr::kable(.) 

```

# HepG2 hepatoma

```{r}

cellline.hepatoma <- (data.EGEOD18494$cell_line == "HepG2 hepatoma")

hepatoma.meanBin <- 
expr.EGEOD18494.hif %>% 
  dplyr::select(c("symbol", data.EGEOD18494$codes[cellline.hepatoma])) %>% meanBinNet(.) 

hepatoma.meanBin %>% 
  knitr::kable(.)
```


```{r}

cellline.rep1 <- (data.EGEOD18494$cell_line == "HepG2 hepatoma" &  data.EGEOD18494$rep == 1)
cellline.rep2 <- (data.EGEOD18494$cell_line == "HepG2 hepatoma" &  data.EGEOD18494$rep == 2)
cellline.rep3 <- (data.EGEOD18494$cell_line == "HepG2 hepatoma" &  data.EGEOD18494$rep == 3)

hepatoma1x <-
expr.EGEOD18494.hif %>%
  dplyr::select(c("symbol", data.EGEOD18494$codes[cellline.rep1]))  %>%
  binNet(.)

# hepatoma1x %>%
#   knitr::kable(.)

hepatoma2x <-
expr.EGEOD18494.hif %>%
  dplyr::select(c("symbol", data.EGEOD18494$codes[cellline.rep2]))  %>%
  binNet(.)

# hepatoma2x %>%
#   knitr::kable(.)

hepatoma3x <-
expr.EGEOD18494.hif %>%
  dplyr::select(c("symbol", data.EGEOD18494$codes[cellline.rep3]))  %>%
  binNet(.)

# hepatoma3x %>%
#   knitr::kable(.)

hepatoma.mean <-
cbind(hepatoma1x,hepatoma2x,hepatoma3x) %>%
  tibble::rownames_to_column('gene') %>%
  mutate_at(vars(-gene), as.numeric) %>%
  mutate(no.ctrl = rowMeans(dplyr::select(.,starts_with("no.control")), na.rm = TRUE)) %>%
  mutate(hy.4h = rowMeans(dplyr::select(.,starts_with("hy.4h")), na.rm = TRUE)) %>%
  mutate(hy.8h = rowMeans(dplyr::select(.,starts_with("hy.8h")), na.rm = TRUE)) %>%
  mutate(hy.12h = rowMeans(dplyr::select(.,starts_with("hy.12h")), na.rm = TRUE)) %>%
  dplyr::select(c("no.ctrl", "hy.4h", "hy.8h", "hy.12h", "gene")) %>%
  mutate_at(c("no.ctrl", "hy.4h", "hy.8h", "hy.12h"), funs(ifelse(. >= 0.5, 1, 0)))  %>%  # consensus with a bies to 1 (>= 0.5)
  tibble::column_to_rownames('gene')

hepatoma.mean %>%  
 knitr::kable(.)
```


# U87 glioma

```{r}

cellline.glioma <- (data.EGEOD18494$cell_line == "U87 glioma")

glioma.meanBin <- 
expr.EGEOD18494.hif %>% 
  dplyr::select(c("symbol", data.EGEOD18494$codes[cellline.glioma])) %>% meanBinNet(.) 

glioma.meanBin %>% 
  knitr::kable(.)
```


```{r}
cellline.rep1 <- (data.EGEOD18494$cell_line == "U87 glioma" &  data.EGEOD18494$rep == 1)
cellline.rep2 <- (data.EGEOD18494$cell_line == "U87 glioma" &  data.EGEOD18494$rep == 2)
cellline.rep3 <- (data.EGEOD18494$cell_line == "U87 glioma" &  data.EGEOD18494$rep == 3)

glioma1x <-
expr.EGEOD18494.hif %>%
  dplyr::select(c("symbol", data.EGEOD18494$codes[cellline.rep1]))  %>%
  binNet(.)

# glioma1x %>%
#   knitr::kable(.)

glioma2x <-
expr.EGEOD18494.hif %>%
  dplyr::select(c("symbol", data.EGEOD18494$codes[cellline.rep2]))  %>%
  binNet(.)

# glioma2x %>%
#   knitr::kable(.)

glioma3x <-
expr.EGEOD18494.hif %>%
  dplyr::select(c("symbol", data.EGEOD18494$codes[cellline.rep3]))  %>%
  binNet(.)

# glioma3x %>%
#   knitr::kable(.)

glioma.mean <-
cbind(glioma1x,glioma2x,glioma3x) %>%
  tibble::rownames_to_column('gene') %>%
  mutate_at(vars(-gene), as.numeric) %>%
  mutate(no.ctrl = rowMeans(dplyr::select(.,starts_with("no.control")), na.rm = TRUE)) %>%
  mutate(hy.4h = rowMeans(dplyr::select(.,starts_with("hy.4h")), na.rm = TRUE)) %>%
  mutate(hy.8h = rowMeans(dplyr::select(.,starts_with("hy.8h")), na.rm = TRUE)) %>%
  mutate(hy.12h = rowMeans(dplyr::select(.,starts_with("hy.12h")), na.rm = TRUE)) %>%
  dplyr::select(c("no.ctrl", "hy.4h", "hy.8h", "hy.12h", "gene")) %>%
  mutate_at(c("no.ctrl", "hy.4h", "hy.8h", "hy.12h"), funs(ifelse(. >= 0.5, 1, 0)))  %>%  # consensus with a bies to 1 (>= 0.5)
  tibble::column_to_rownames('gene')

glioma.mean %>%  
 knitr::kable(.)

```


<!-- ```{r  include=FALSE} -->
<!-- breast1x.net <- reconstructNetwork(breast1x, method="bestfit", returnPBN=TRUE, readableFunctions=TRUE) -->
<!-- breast2x.net <- reconstructNetwork(breast2x, method="bestfit", returnPBN=TRUE, readableFunctions=TRUE) -->
<!-- breast3x.net <- reconstructNetwork(breast3x, method="bestfit", returnPBN=TRUE, readableFunctions=TRUE) -->

<!-- hepatoma1x.net <- reconstructNetwork(hepatoma1x, method="bestfit", returnPBN=TRUE, readableFunctions=TRUE) -->
<!-- hepatoma2x.net <- reconstructNetwork(hepatoma2x, method="bestfit", returnPBN=TRUE, readableFunctions=TRUE) -->
<!-- hepatoma3x.net <- reconstructNetwork(hepatoma3x, method="bestfit", returnPBN=TRUE, readableFunctions=TRUE) -->

<!-- glioma1x.net <- reconstructNetwork(glioma1x, method="bestfit", returnPBN=TRUE, readableFunctions=TRUE) -->
<!-- glioma2x.net <- reconstructNetwork(glioma2x, method="bestfit", returnPBN=TRUE, readableFunctions=TRUE) -->
<!-- glioma3x.net <- reconstructNetwork(glioma3x, method="bestfit", returnPBN=TRUE, readableFunctions=TRUE) -->

<!-- breast1x.p <- plotNetworkWiring(breast1x.net, plotIt=F) -->
<!-- breast2x.p <- plotNetworkWiring(breast2x.net, plotIt=F) -->
<!-- breast3x.p <- plotNetworkWiring(breast3x.net, plotIt=F) -->

<!-- hepatoma1x.p <- plotNetworkWiring(hepatoma1x.net, plotIt=F) -->
<!-- hepatoma2x.p <- plotNetworkWiring(hepatoma2x.net, plotIt=F) -->
<!-- hepatoma3x.p <- plotNetworkWiring(hepatoma3x.net, plotIt=F) -->

<!-- glioma1x.p <- plotNetworkWiring(glioma1x.net, plotIt=F) -->
<!-- glioma2x.p <- plotNetworkWiring(glioma2x.net, plotIt=F) -->
<!-- glioma3x.p <- plotNetworkWiring(glioma3x.net, plotIt=F) -->

<!-- # All breast cancer nets merged: -->
<!-- breast.all <- reconstructNetwork(list(breast1x, breast2x, breast3x), -->
<!--                                method="bestfit",returnPBN=TRUE,readableFunctions=TRUE) -->
<!-- breast.all.p <- plotNetworkWiring(breast.all, plotIt=F) -->

<!-- # All hepatoma cancer nets merged: -->
<!-- hepatoma.all <- reconstructNetwork(list(hepatoma1x, hepatoma2x, hepatoma3x), -->
<!--                                method="bestfit",returnPBN=TRUE,readableFunctions=TRUE) -->
<!-- hepatoma.all.p <- plotNetworkWiring(hepatoma.all, plotIt=F) -->

<!-- # All hepatoma cancer nets merged: -->
<!-- glioma.all <- reconstructNetwork(list(glioma1x, glioma2x, glioma3x), -->
<!--                                method="bestfit",returnPBN=TRUE,readableFunctions=TRUE) -->
<!-- glioma.all.p <- plotNetworkWiring(glioma.all, plotIt=F) -->


<!-- # Mean AFTER binarize the replicates of breast cancer net : -->
<!-- mean.net <- reconstructNetwork(breast.mean, -->
<!--                                method="bestfit",returnPBN=TRUE,readableFunctions=TRUE) -->

<!-- mean.p <- plotNetworkWiring(mean.net, plotIt=F) -->
<!-- ``` -->


# Mean AFTER binarize the replicates of breast cancer net :

## MDA-MB231 breast cancer

```{r}
breast.mean.net <- reconstructNetwork(breast.mean,
                               method="bestfit", returnPBN=TRUE,readableFunctions=TRUE)

breast.mean.p <- plotNetworkWiring(breast.mean.net, plotIt=F)


```

## HepG2 hepatoma

```{r}
hepatoma.mean.net <- reconstructNetwork(hepatoma.mean,
                               method="bestfit", returnPBN=TRUE,readableFunctions=TRUE)

hepatoma.mean.p <- plotNetworkWiring(hepatoma.mean.net, plotIt=F)

```

## U87 glioma

```{r}
glioma.mean.net <- reconstructNetwork(glioma.mean,
                               method="bestfit", returnPBN=TRUE,readableFunctions=TRUE)

glioma.mean.p <- plotNetworkWiring(glioma.mean.net, plotIt=F)
```

# Mean BEFORE binarize the replicates :

## MDA-MB231 breast cancer

```{r}
breast.meanBin.net <- reconstructNetwork(breast.meanBin,
                               method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)

breast.meanBin.p <- plotNetworkWiring(breast.meanBin.net, plotIt=F)
```

## HepG2 hepatoma

```{r}
hepatoma.meanBin.net <- reconstructNetwork(hepatoma.meanBin,
                               method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)

hepatoma.meanBin.p <- plotNetworkWiring(hepatoma.meanBin.net, plotIt=F)
```

## U87 glioma

```{r}
glioma.meanBin.net <- reconstructNetwork(glioma.meanBin,
                               method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)

glioma.meanBin.p <- plotNetworkWiring(glioma.meanBin.net, plotIt=F)
```


# Network inference:

<!-- ```{r } -->

<!-- # MDA-MB231 breast cancer - 4 time-points -->
<!-- par(mfrow = c(1,3)) -->
<!-- plot(breast1x.p, vertex.label.color="#440154ff", vertex.color="lightblue", vertex.frame.color="white", layout=layout_in_circle, edge.curved=.3, -->
<!--      main="MDA-MB231 breast\n 4 time-points, replicate 1") -->
<!-- plot(breast2x.p, vertex.label.color="#440154ff", vertex.color="lightblue", vertex.frame.color="white", layout=layout_in_circle, edge.curved=.3, -->
<!--      main="MDA-MB231 breast\n 4 time-points, replicate 2") -->
<!-- plot(breast3x.p, vertex.label.color="#440154ff", vertex.color="lightblue", vertex.frame.color="white", layout=layout_in_circle, edge.curved=.3, -->
<!--      main="MDA-MB231 breast\n 4 time-points, replicate 3") -->

<!-- par(mfrow = c(1,1)) -->
<!-- plot(breast.all.p, vertex.label.color="#440154ff", vertex.color="lightblue", vertex.frame.color="white", layout=layout_in_circle, edge.curved=.3, -->
<!--      main="MDA-MB231 breast\n 4 time-points, all replicate") -->

<!-- # HepG2 hepatoma -->
<!-- par(mfrow = c(1,3)) -->
<!-- plot(hepatoma1x.p, vertex.label.color="#440154ff", vertex.color="lightblue", vertex.frame.color="white", layout=layout_in_circle, edge.curved=.3, -->
<!--      main="HepG2 hepatoma\n 4 steps, replicate 1") -->
<!-- plot(hepatoma2x.p, vertex.label.color="#440154ff", vertex.color="lightblue", vertex.frame.color="white", layout=layout_in_circle, edge.curved=.3, -->
<!--      main="HepG2 hepatoma\n 4 steps, replicate 2") -->
<!-- plot(hepatoma3x.p, vertex.label.color="#440154ff", vertex.color="lightblue", vertex.frame.color="white", layout=layout_in_circle, edge.curved=.3, -->
<!--      main="HepG2 hepatoma\n 4 steps, replicate 3") -->

<!-- par(mfrow = c(1,1)) -->
<!-- plot(hepatoma.all.p, vertex.label.color="#440154ff", vertex.color="lightblue", vertex.frame.color="white", layout=layout_in_circle, edge.curved=.3, -->
<!--      main="HepG2 hepatoma\n 4 steps, replicate 3") -->

<!-- # U87 glioma -->
<!-- par(mfrow = c(1,3)) -->
<!-- plot(glioma1x.p, vertex.label.color="#440154ff", vertex.color="lightblue", vertex.frame.color="white", layout=layout_in_circle, edge.curved=.3, -->
<!--      main="U87 glioma\n 4 steps, replicate 1") -->
<!-- plot(glioma2x.p, vertex.label.color="#440154ff", vertex.color="lightblue", vertex.frame.color="white", layout=layout_in_circle, edge.curved=.3, -->
<!--      main="U87 glioma\n 4 steps, replicate 2") -->
<!-- plot(glioma3x.p, vertex.label.color="#440154ff", vertex.color="lightblue", vertex.frame.color="white", layout=layout_in_circle, edge.curved=.3, -->
<!--      main="U87 glioma\n 4 steps, replicate 3") -->

<!-- par(mfrow = c(1,1)) -->
<!-- plot(glioma.all.p, vertex.label.color="#440154ff", vertex.color="lightblue", vertex.frame.color="white", layout=layout_in_circle, edge.curved=.3, -->
<!--      main="U87 glioma\n 4 steps, replicate 3") -->

<!-- ``` -->


# Mean BEFORE binarize the replicates of breast cancer net :

```{r}
par(mfrow = c(1,1))
plot(breast.meanBin.p, vertex.label.color="#440154ff", vertex.color="lightblue", vertex.frame.color="white", layout=layout_in_circle, edge.curved=.3,
     main="MDA-MB231 breast\n 4 time-points, Mean binarized replicates")
```

```{r}
print(breast.meanBin.net)
```

```{r}

try({
sink("../data/ATOTS_inferred_EGEOD18494_breast.bn")
cat("targets, factors\n")
cat("EP300, (!TP53 | HIF1A) \n")
cat("HIF1A,  !EP300 \n")
cat("MDM2, (TP53 | !HIF1A)\n")
cat("TP53, EP300\n")
cat("VHL, (!TP53 | HIF1A)\n")
sink()}, silent = T)
```

```{r}
net <- loadNetwork("../data/ATOTS_inferred_EGEOD18494_breast.bn")
print(net)
```

```{r}

attr.syn <- getAttractors(net, type = "synchronous")
plotAttractors(attr.syn, title = "MDA-MB231 breast, mean Before binarize replicates")
```

# Mean AFTER binarize the replicates of breast cancer net :

```{r}
par(mfrow = c(1,1))
plot(breast.mean.p, vertex.label.color="#440154ff", vertex.color="lightblue", vertex.frame.color="white", layout=layout_in_circle, edge.curved=.3,
     main="MDA-MB231 breast\n 4 time-points, Mean After Binarize replicates")
```

```{r}
print(breast.mean.net)
```

```{r}

try({
sink("../data/ATOTS_inferred_EGEOD18494_breast_meanAfter.bn")
cat("targets, factors\n")
cat("EP300, 1\n")
cat("HIF1A, (!TP53 | !EP300 )\n")
cat("MDM2, (TP53 | EP300)\n")
cat("TP53, 1\n")
cat("VHL, HIF1A\n")
sink()}, silent = T)
```


```{r}
net <- loadNetwork("../data/ATOTS_inferred_EGEOD18494_breast_meanAfter.bn")
print(net)
```

```{r}
attr.syn <- getAttractors(net, type = "synchronous")
plotAttractors(attr.syn, title = "MDA-MB231 breast, mean After binarize replicates")
```

# Mean BEFORE binarize the replicates of hepatoma cancer net :

```{r}
par(mfrow = c(1,1))
plot(hepatoma.meanBin.p, vertex.label.color="#440154ff", vertex.color="lightblue", vertex.frame.color="white", layout=layout_in_circle, edge.curved=.3,
     main="HepG2 hepatoma\n 4 time-points, Mean Before binarize replicates")
```

```{r}
print(hepatoma.meanBin.net)
```

```{r}

try({
sink("../data/ATOTS_inferred_EGEOD18494_hepatoma.bn")
cat("targets, factors\n")
cat("EP300, (!MDM2 | !EP300) \n")
cat("HIF1A,  (MDM2 | EP300) \n")
cat("MDM2, (!MDM2 | EP300)\n")
cat("TP53, VHL\n")
cat("VHL, 0\n")
sink()}, silent = T)
```

```{r}
net <- loadNetwork("../data/ATOTS_inferred_EGEOD18494_hepatoma.bn")
print(net)
```

```{r}

attr.syn <- getAttractors(net, type = "synchronous")
plotAttractors(attr.syn, title = "HepG2 hepatoma, mean Before binarize replicates")
```

# Mean AFTER binarize the replicates of HepG2 hepatoma cancer net :

```{r}
par(mfrow = c(1,1))
plot(hepatoma.mean.p, vertex.label.color="#440154ff", vertex.color="lightblue", vertex.frame.color="white", layout=layout_in_circle, edge.curved=.3,
     main="HepG2 hepatoma\n 4 time-points, Mean After Binarize replicates")
```

```{r}
print(hepatoma.mean.net)
```

```{r}

try({
sink("../data/ATOTS_inferred_EGEOD18494_hepatoma_meanAfter.bn")
cat("targets, factors\n")
cat("EP300, (VHL | !MDM2) | !EP300 \n")
cat("HIF1A, (!VHL | MDM2 ) | EP300\n")
cat("MDM2, (VHL | !MDM2) | !EP300\n")
cat("TP53, 1\n")
cat("VHL, (!VHL | MDM2) | EP300\n")
sink()}, silent = T)
```

```{r}
net <- loadNetwork("../data/ATOTS_inferred_EGEOD18494_hepatoma_meanAfter.bn")
print(net)
```

```{r}

attr.syn <- getAttractors(net, type = "synchronous")
plotAttractors(attr.syn, title = "HepG2 hepatoma, mean After binarize replicates")
```

# Mean BEFORE binarize the replicates of U87 glioma cancer net :

```{r}
par(mfrow = c(1,1))
plot(glioma.meanBin.p, vertex.label.color="#440154ff", vertex.color="lightblue", vertex.frame.color="white", layout=layout_in_circle, edge.curved=.3,
     main="U87 glioma\n 4 time-points, Mean binarized replicates")
```

```{r}
print(glioma.meanBin.net)
```

```{r}

try({
sink("../data/ATOTS_inferred_EGEOD18494_glioma.bn")
cat("targets, factors\n")
cat("EP300, HIF1A \n")
cat("HIF1A,  MDM2 \n")
cat("MDM2, 0\n")
cat("TP53, !TP53\n")
cat("VHL, HIF1A\n")
sink()}, silent = T)
```

```{r}
net <- loadNetwork("../data/ATOTS_inferred_EGEOD18494_glioma.bn")
print(net)
```

```{r}

attr.syn <- getAttractors(net, type = "synchronous")
plotAttractors(attr.syn, title = "U87 glioma, mean Before binarize replicates")
```

# Mean AFTER binarize the replicates of U87 glioma cancer net :

```{r}
par(mfrow = c(1,1))
plot(glioma.mean.p, vertex.label.color="#440154ff", vertex.color="lightblue", vertex.frame.color="white", layout=layout_in_circle, edge.curved=.3,
     main="U87 glioma\n 4 time-points, Mean After Binarize replicates")
```

```{r}
print(glioma.mean.net)
```

```{r}

try({
sink("../data/ATOTS_inferred_EGEOD18494_glioma_meanAfter.bn")
cat("targets, factors\n")
cat("EP300, (!!TP53 | !EP300)\n")
cat("HIF1A, MDM2\n")
cat("MDM2, 0\n")
cat("TP53, !MDM2\n")
cat("VHL, HIF1A\n")
sink()}, silent = T)
```


```{r}
net <- loadNetwork("../data/ATOTS_inferred_EGEOD18494_hepatoma_meanAfter.bn")
print(net)
```

```{r}

attr.syn <- getAttractors(net, type = "synchronous")
plotAttractors(attr.syn, title = "U87 glioma, Mean After Binarize replicates")
```
