---
title: "BoolNet Inference MCF-7 breast (GSE47533)"
output:
  pdf_document: default
  github_document: 
    df_print: paged
    html_preview: FALSE
    keep_html: FALSE
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_format = "all") })    
---

```{r error=TRUE, message=FALSE, warning=FALSE, include=FALSE, purl=FALSE, results='hide'}
## This chunk automatically generates a text .R version of this script when running within knitr.
# input  = knitr::current_input()  # filename of input document
# output = paste(tools::file_path_sans_ext(input), 'R', sep = '.')
# try(knitr::purl(input,output,documentation=2,quiet=T), silent = TRUE)
# # Avoid duplicate label error of knitr::purl
options(knitr.duplicate.label = 'allow')
# Code to browse the markdown file with renderized images.
knitr::opts_chunk$set(
  fig.path = "figs/GSE47533-"
)
```

Integrated analysis of microRNA and mRNA expression and association with HIF binding in MCF-7 cells under hypoxia (GSE47533)

Camps C, Saini HK, Mole DR, Choudhry H et al. Integrated analysis of microRNA and mRNA expression and association with HIF binding reveals the complexity of microRNA expression regulation under hypoxia. Mol Cancer 2014 Feb 11;13:28. PMID: 24517586

https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE47533

This SuperSeries is composed of the following SubSeries:

GSE47532 	MCF-7 cells under hypoxia [miRNA] - Samples (11) - 822 miRNA

GSE47533 	MCF-7 cells under hypoxia [mRNA] - GPL6884 - Samples (12)

GSE47602 	MCF-7 cells under hypoxia (miRNA-Seq) - Samples (8) - missing


```{r message=FALSE, warning=FALSE}
packages_cran = c("igraph", "BoolNet", "BiocManager", "tidyverse", "fs", "ff", "effectsize")
# Install and load packages
package.check <- lapply(packages_cran, FUN = function(x) {
  if (!require(x, character.only = TRUE)) {
    install.packages(x, dependencies = TRUE)
    library(x, character.only = TRUE)
  }
})
# For oligo and ArrayExpress First install:
#install.packages('https://cran.r-project.org/src/contrib/Archive/ff/ff_2.2-14.tar.gz',repos=NULL)
# packages_bioconductor = c("Biobase", "GEOquery", "affyPLM", "ArrayExpress", "illuminaHumanv3.db")
# # Install and load packages
# package.check <- lapply(packages_bioconductor, FUN = function(x) {
#   if (!require(x, character.only = TRUE)) {
#     BiocManager::install(x, dependencies = TRUE)
#     library(x, character.only = TRUE)
#   }
# })
rm(package.check, packages_cran)
```

<!-- ```{r} -->

<!-- download_dir <- fs::path(".data_tmp") -->
<!-- if (!dir_exists(download_dir)) { dir_create(download_dir) } -->

<!-- if (!dir_exists(".data_tmp/GSE47533_series_matrix.txt.gz")) { -->
<!--   GSE47533 <-getGEO("GSE47533", destdir = download_dir, GSEMatrix = T) -->
<!--   } else{ -->
<!--   GSE47533 <-getGEO(filename=".data_tmp/GSE47533_series_matrix.txt.gz") -->
<!-- } -->

<!-- # Normalisation: the data is already normalised -->
<!-- expr.GSE47533 <- as.matrix(exprs(normalize.ExpressionSet.quantiles(GSE47533[[1]]))) -->
<!-- prob.GSE47533 <- unique(rownames(expr.GSE47533)) -->
<!-- data.GSE47533 <- pData(GSE47533[[1]]) -->

<!-- data.GSE47533 <- data.frame( -->
<!--                   codes = as.character(data.GSE47533$geo_accession), -->
<!--                   cell_line = "MCF7", -->
<!--                   time = data.GSE47533$`time of exposure:ch1`, -->
<!--                   condition = substr(as.character(data.GSE47533$description), 1, 4), -->
<!--                   rep = data.GSE47533$description.1) -->

<!-- data.GSE47533 <- data.GSE47533 %>% -->
<!--   mutate(rep = recode(rep, "replicate 1" = 1, -->
<!--                            "replicate 2" = 2, -->
<!--                            "replicate 3" = 3)) %>% -->
<!--   mutate_at(vars(time), as.character)  %>% -->
<!--   mutate(time = ifelse(condition == "Norm", 0, time)) -->

<!-- # Convert the probes to Symbol names -->
<!-- # The below function call will return a datafram with probe_id, gene symbol -->
<!-- # and ŕefgene_id for your data -->
<!-- anno.GSE47533 <- AnnotationDbi::select(illuminaHumanv3.db, -->
<!--        keys = prob.GSE47533, -->
<!--        columns=c("ENSEMBL", "SYMBOL", "GENENAME"), -->
<!--        keytype="PROBEID") -->

<!-- colnames(anno.GSE47533) <- c("probes", "ensgene", "symbol", "description") -->

<!-- rm(download_dir, GSE47533,  prob.GSE47533) -->

<!-- save.image("../data/data.GSE47533.Rdata") -->
<!-- ``` -->


# Load the pre-processed data

```{r}
load("../data/data.GSE47533.Rdata")
cols <- colnames(expr.GSE47533)
rows <- rownames(expr.GSE47533)
expr.GSE47533 <- as.data.frame(matrix(effectsize::normalize(as.matrix(expr.GSE47533)), ncol = length(cols), nrow = length(rows) ))
colnames(expr.GSE47533) <- cols
rownames(expr.GSE47533) <- rows 
```

# Selecting the HIF Genes 
    
```{r}
# Selected genes from HIF Axis
hif.symbols <- c("TP53", "HIF1A", "EP300", "MDM2", "VHL")
hif.probes <- anno.GSE47533$probes[anno.GSE47533$symbol %in% hif.symbols]

# Select the probes and genes
expr.GSE47533.hif <- data.frame(expr.GSE47533) %>% 
  rownames_to_column('probes') %>% 
  filter(probes %in% hif.probes) %>% 
  merge(anno.GSE47533[anno.GSE47533$symbol %in% hif.symbols, c("probes","symbol")], by = "probes") %>% 
  #distinct(symbol, .keep_all = TRUE) %>% # Take the first one
  dplyr::select(!(probes)) %>% 
  arrange(symbol)
```

# Example of Binarizing

```{r}
cols <- (data.GSE47533$rep == 1)
breast1_MCF7 <- 
expr.GSE47533.hif %>% 
  dplyr::select(c("symbol", data.GSE47533$codes[cols])) %>% arrange(symbol) %>% 
  arrange(symbol) %>% 
  rename_at(vars(data.GSE47533$codes[cols]),
            ~paste0(substr(data.GSE47533$condition[cols],1,2),".",
                    data.GSE47533$time[cols],".",
                    substr(data.GSE47533$cell_line[cols],1,2)))

knitr::kable(breast1_MCF7)

binarizeTimeSeries(breast1_MCF7[,-1], method="kmeans")$binarizedMeasurements  %>% 
  data.frame(.)  %>% 
  add_column(symbol = breast1_MCF7$symbol, .before=0) %>% 
  knitr::kable(.)
  
binarizeTimeSeries(breast1_MCF7[,-1], method="kmeans")$binarizedMeasurements  %>% 
  data.frame(.)  %>% 
  aggregate(., list(symbol = breast1_MCF7$symbol), mean) %>% 
  mutate_at(vars(-symbol), funs(ifelse(. >= 0.5, 1, 0))) %>% 
  #rbind(., c("O2", 1,0,0,0)) %>%  # removing O2
  knitr::kable(.)
```



```{r}

# Function to binarize according an consensus mean of probes, add the O2 state and rename columns 
binNet <- function(b){
  
  cols <- data.GSE47533$codes %in% names(b)
  
  binarizeTimeSeries(b[,-1], method="kmeans")$binarizedMeasurements  %>% 
  as.data.frame(.)  %>% 
  aggregate(., list(symbol = b$symbol), mean) %>%  # mean of binarized probes
  mutate_at(vars(-symbol), funs(ifelse(. >= 0.5, 1, 0))) %>%  # consensus with a bies to 1 (>= 0.5)
  #rbind(., c("O2", 1,0,0,0)) %>%  # removing O2
    rename_at(vars(data.GSE47533$codes[cols] ),
            ~paste0(substr(data.GSE47533$condition[cols],1,2),".",
                    data.GSE47533$time[cols],".",
                    substr(data.GSE47533$cell_line[cols],1,2), ".",
                    data.GSE47533$rep[cols])) %>% 
  column_to_rownames("symbol")
  
}
```


```{r}
# Function to calculate the mean and binarize after, according an consensus mean of probes 

meanBinNet <- function(b){
  
  cols <- data.GSE47533$codes %in% names(b)
  
  b <-b %>% 
  rename_at(vars(data.GSE47533$codes[cols] ),
            ~paste0(substr(data.GSE47533$condition[cols],1,2),".",
                    data.GSE47533$time[cols],".",
                    substr(data.GSE47533$cell_line[cols],1,2), ".",
                    data.GSE47533$rep[cols])) %>% 
  mutate(No.0.MC = rowMeans(dplyr::select(.,starts_with("No.0.MC")), na.rm = TRUE)) %>% 
  mutate(Hy.16h.MC = rowMeans(dplyr::select(.,starts_with("Hy.16h.MC")), na.rm = TRUE)) %>% 
  mutate(Hy.32h.MC = rowMeans(dplyr::select(.,starts_with("Hy.32h.MC")), na.rm = TRUE)) %>% 
  mutate(Hy.48h.MC = rowMeans(dplyr::select(.,starts_with("Hy.48h.MC")), na.rm = TRUE)) %>%
  dplyr::select(c("symbol", "No.0.MC", "Hy.16h.MC", "Hy.32h.MC", "Hy.48h.MC")) 
  
  binarizeTimeSeries(b[,-1], method="kmeans")$binarizedMeasurements  %>% 
  as.data.frame(.)  %>% 
  aggregate(., list(symbol = b$symbol), mean) %>%  # mean of binarized probes
  mutate_at(vars(-symbol), funs(ifelse(. >= 0.5, 1, 0))) %>%  # consensus with a bies to 1 (>= 0.5)
  #rbind(., c("O2", 1,0,0,0)) %>% 
  column_to_rownames("symbol")
  
}

```


```{r}

breast_MCF7.1 <- 
expr.GSE47533.hif %>% 
  dplyr::select(c("symbol", data.GSE47533$codes[data.GSE47533$rep == 1]))  %>% 
  binNet(.) 

breast_MCF7.1 %>% 
  knitr::kable(.)

breast_MCF7.2 <- 
expr.GSE47533.hif %>% 
  dplyr::select(c("symbol", data.GSE47533$codes[data.GSE47533$rep == 2]))  %>% 
  binNet(.) 

breast_MCF7.2  %>% 
  knitr::kable(.)

breast_MCF7.3 <- 
expr.GSE47533.hif %>% 
  dplyr::select(c("symbol", data.GSE47533$codes[data.GSE47533$rep == 3]))  %>% 
  binNet(.) 

breast_MCF7.3 %>% 
  knitr::kable(.)

breast_MCF7.mean <- 
cbind(breast_MCF7.1, breast_MCF7.2,breast_MCF7.3) %>%
  tibble::rownames_to_column('gene') %>%
  mutate_at(vars(-gene), as.numeric) %>% 
  mutate(No.0.MC = rowMeans(dplyr::select(.,starts_with("No.0.MC")), na.rm = TRUE)) %>% 
  mutate(Hy.16h.MC = rowMeans(dplyr::select(.,starts_with("Hy.16h.MC")), na.rm = TRUE)) %>% 
  mutate(Hy.32h.MC = rowMeans(dplyr::select(.,starts_with("Hy.32h.MC")), na.rm = TRUE)) %>% 
  mutate(Hy.48h.MC = rowMeans(dplyr::select(.,starts_with("Hy.48h.MC")), na.rm = TRUE)) %>%
  dplyr::select(c("No.0.MC", "Hy.16h.MC", "Hy.32h.MC", "Hy.48h.MC", "gene")) %>%
  mutate_at(c("No.0.MC", "Hy.16h.MC", "Hy.32h.MC", "Hy.48h.MC"), funs(ifelse(. >= 0.5, 1, 0)))  %>%  # consensus with a bies to 1 (>= 0.5)  
  tibble::column_to_rownames('gene')

breast_MCF7.mean %>% 
  knitr::kable(.)

breast.meanBin <- 
expr.GSE47533.hif %>% 
  dplyr::select(c("symbol", data.GSE47533$codes)) %>% meanBinNet(.) 

breast.meanBin %>% 
  knitr::kable(.)

# All breast cancer nets merged:
all.nets <- reconstructNetwork(list(breast_MCF7.1, breast_MCF7.2, breast_MCF7.3),
                               method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)

all.p <- plotNetworkWiring(all.nets, plotIt=F)




# # Mean of replicate breast cancer net without O2 :
# mean.nox.net <- reconstructNetwork(breast_MCF7.mean[c(1:5),],
#                                method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)
# 
# mean.nox.p <- plotNetworkWiring(mean.nox.net, plotIt=F)

# Mean of replicate breast cancer net with only Hypoxia state:
# 
# mean.hyp.net <- reconstructNetwork(breast_MCF7.mean[,c("Hy.16h.MC", "Hy.32h.MC", "Hy.48h.MC")],
#                                method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)
# 
# mean.hyp.p <- plotNetworkWiring(mean.hyp.net, plotIt=F)

  
# Mean AFTER binarize the replicates of breast cancer net :

mean.net <- reconstructNetwork(breast_MCF7.mean,
                               method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)

mean.p <- plotNetworkWiring(mean.net, plotIt=F)


# Mean BEFORE binarize the replicates of breast cancer net :

meanBin.net <- reconstructNetwork(breast.meanBin,
                               method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)

meanBin.p <- plotNetworkWiring(meanBin.net, plotIt=F)

```


# MCF7 breast cancer

```{r}
# MCF7 breast cancer - 4 time-points
breast_MCF7.1.net <- reconstructNetwork(breast_MCF7.1, method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)
breast_MCF7.2.net <- reconstructNetwork(breast_MCF7.2, method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)
breast_MCF7.3.net <- reconstructNetwork(breast_MCF7.3, method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)

breast_MCF7.1.p <- plotNetworkWiring(breast_MCF7.1.net, plotIt=F)
breast_MCF7.2.p <- plotNetworkWiring(breast_MCF7.2.net, plotIt=F)
breast_MCF7.3.p <- plotNetworkWiring(breast_MCF7.3.net, plotIt=F)
```

<!-- ```{r} -->
<!-- # MCF7 breast - 4 steps, replicate 1 -->
<!-- print(breast_MCF7.1.net) -->

<!-- # MCF7 breast - 4 steps, replicate 2 -->
<!-- print(breast_MCF7.2.net) -->

<!-- # MCF7 breast - 4 steps, replicate 3 -->
<!-- print(breast_MCF7.3.net) -->

<!-- # MCF7 breast - 4 steps, all replicates -->
<!-- print(all.nets) -->

<!-- # MCF7 breast - 4 steps, all replicates -->
<!-- print(mean.net) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- par(mfrow = c(1,1)) -->
<!-- plot(all.p, vertex.label.color="#440154ff", vertex.color="lightblue", vertex.frame.color="white", layout=layout_in_circle, edge.curved=.3, -->
<!--      main="MCF7 breast\n 4 time-points, all replicates") -->
<!-- ``` -->


<!-- ```{r} -->
<!-- par(mfrow = c(1,1)) -->
<!-- plot(mean.hyp.p, vertex.label.color="#440154ff", vertex.color="lightblue", vertex.frame.color="white", layout=layout_in_circle,# edge.curved=.3, -->
<!--      main="MCF7 breast\n 3 time-points, without Normoxia time-point") -->
<!-- ``` -->

# Mean AFTER binarize the replicates of breast cancer net :

```{r}
# MCF7 breast cancer - 4 time-points 
par(mfrow = c(1,3))
plot(breast_MCF7.1.p, vertex.label.color="#440154ff", vertex.color="lightblue", vertex.frame.color="white", layout=layout_in_circle, edge.curved=.3,
     main="MCF7 breast\n 4 time-points, replicate 1")
plot(breast_MCF7.2.p, vertex.label.color="#440154ff", vertex.color="lightblue", vertex.frame.color="white", layout=layout_in_circle, edge.curved=.3,
     main="MCF7 breast\n 4 time-points, replicate 2")
plot(breast_MCF7.3.p, vertex.label.color="#440154ff", vertex.color="lightblue", vertex.frame.color="white", layout=layout_in_circle, edge.curved=.3,
     main="MCF7 breast\n 4 time-points, replicate 3")
```


```{r}
par(mfrow = c(1,1))
plot(mean.p, vertex.label.color="#440154ff", vertex.color="lightblue", vertex.frame.color="white", layout=layout_in_circle, edge.curved=.3,
     main="MCF7 breast\n 4 time-points, Mean replicates")
```

```{r}
print(mean.net)
```

<!-- ```{r} -->
<!-- par(mfrow = c(1,1)) -->
<!-- plot(mean.nox.p, vertex.label.color="#440154ff", vertex.color="lightblue", vertex.frame.color="white", layout=layout_in_circle, edge.curved=.3, -->
<!--      main="MCF7 breast\n 4 time-points, without O2") -->
<!-- ``` -->


```{r results='hide'}

# sink("ATOTS_inferred.bn")
# cat("targets, factors\n")
# cat("EP300, (TP53 & !VHL) | (!TP53 & VHL) | (TP53 & !VHL) | (!HIF1A & TP53) |  (!HIF1A & TP53) | (HIF1A & !TP53) | (!EP300 & !VHL) | (!EP300 & !VHL) | (EP300 & VHL) | (!EP300 & !HIF1A) | (EP300 & HIF1A)\n")
# cat("HIF1A, !TP53 | EP300\n")
# cat("MDM2, TP53 & !VHL\n")
# cat("TP53, VHL | HIF1A\n")
# cat("VHL, !TP53 | !EP300\n")
# sink()
try({
sink("../data/ATOTS_inferred_GSE47533.bn")
cat("targets, factors\n")
cat("EP300, (!HIF1A & TP53) | (HIF1A & !TP53)\n")
cat("HIF1A, EP300\n")
cat("MDM2, EP300\n")
cat("TP53, HIF1A\n")
cat("VHL, !TP53\n")
sink()}, silent = T)
```

```{r}
net <- loadNetwork("../data/ATOTS_inferred_GSE47533.bn")
print(net)
```


```{r}

attr.syn <- getAttractors(net, type = "synchronous")
plotAttractors(attr.syn)
```



# Mean BEFORE binarize the replicates of breast cancer net :

```{r}
par(mfrow = c(1,1))
plot(meanBin.p, vertex.label.color="#440154ff", vertex.color="lightblue", vertex.frame.color="white", layout=layout_in_circle, edge.curved=.3,
     main="MDA-MB231 breast\n 4 time-points, Mean BEFORE binarize replicates")
```

```{r}
print(meanBin.net)
```

<!-- ```{r} -->

<!-- try({ -->
<!-- sink("../data/ATOTS_inferred_GSE47533_meanBin.bn") -->
<!-- cat("targets, factors\n") -->
<!-- cat("EP300, (!TP53 | HIF1A) \n") -->
<!-- cat("HIF1A,  !EP300 \n") -->
<!-- cat("MDM2, (TP53 | !HIF1A)\n") -->
<!-- cat("TP53, EP300\n") -->
<!-- cat("VHL, (!TP53 | HIF1A)\n") -->
<!-- sink()}, silent = T) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- net <- loadNetwork("../data/ATOTS_inferred_GSE47533_meanBin.bn") -->
<!-- print(net) -->
<!-- ``` -->


<!-- ```{r} -->

<!-- attr.syn <- getAttractors(net, type = "synchronous") -->
<!-- plotAttractors(attr.syn) -->
<!-- ``` -->