---
title: "BoolNet Inference (GSE47533)"
output:
  github_document: 
    df_print: paged
    html_preview: FALSE
    keep_html: TRUE
  html_notebook: default
  pdf_document: default
---

```{r error=TRUE, message=FALSE, warning=FALSE, include=FALSE, purl=FALSE, results='hide'}
## This chunk automatically generates a text .R version of this script when running within knitr.
input  = knitr::current_input()  # filename of input document
output = paste(tools::file_path_sans_ext(input), 'R', sep = '.')
try(knitr::purl(input,output,documentation=2,quiet=T), silent = TRUE)
# Avoid duplicate label error of knitr::purl
options(knitr.duplicate.label = 'allow')
# Code to browse the markdown file with renderized images.
knitr::opts_chunk$set(
  fig.path = "figs/GSE47533-"
)
```

Integrated analysis of microRNA and mRNA expression and association with HIF binding in MCF-7 cells under hypoxia (GSE47533)

Camps C, Saini HK, Mole DR, Choudhry H et al. Integrated analysis of microRNA and mRNA expression and association with HIF binding reveals the complexity of microRNA expression regulation under hypoxia. Mol Cancer 2014 Feb 11;13:28. PMID: 24517586

https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE47533

Here the non-normalised data:
https://ftp.ncbi.nlm.nih.gov/geo/series/GSE47nnn/GSE47533/suppl/GSE47533%5Fnon%2Dnormalized%5Fdata%2Etxt%2Egz

This SuperSeries is composed of the following SubSeries:

GSE47532 	MCF-7 cells under hypoxia [miRNA] - GPL8227 Agilent-019118 Human miRNA Microarray 2.0 G4470B - Samples (11) - 822 miRNA

GSE47533 	MCF-7 cells under hypoxia [mRNA] - GPL6884 Illumina HumanWG-6 v3.0 expression beadchip - Samples (12)

GSE47602 	MCF-7 cells under hypoxia (miRNA-Seq) - GPL11154 Illumina HiSeq 2000 (Homo sapiens) - Samples (8) - Don't exist


```{r message=FALSE, warning=FALSE}
packages_cran = c("igraph", "BoolNet", "BiocManager", "tidyverse", "fs", "ff")

# Install and load packages
package.check <- lapply(packages_cran, FUN = function(x) {
  if (!require(x, character.only = TRUE)) {
    install.packages(x, dependencies = TRUE)
    library(x, character.only = TRUE)
  }
})

# For oligo and ArrayExpress First install:
#install.packages('https://cran.r-project.org/src/contrib/Archive/ff/ff_2.2-14.tar.gz',repos=NULL)

packages_bioconductor = c("Biobase", "GEOquery", "affyPLM", "ArrayExpress", "illuminaHumanv3.db")

# Install and load packages
package.check <- lapply(packages_bioconductor, FUN = function(x) {
  if (!require(x, character.only = TRUE)) {
    BiocManager::install(x, dependencies = TRUE)
    library(x, character.only = TRUE)
  }
})

rm(package.check, packages_bioconductor, packages_cran)

```

<!-- ```{r} -->

<!-- download_dir <- fs::path(".data_tmp") -->
<!-- if (!dir_exists(download_dir)) { dir_create(download_dir) } -->

<!-- if (!dir_exists(".data_tmp/GSE47533_series_matrix.txt.gz")) { -->
<!--   GSE47533 <-getGEO("GSE47533", destdir = download_dir, GSEMatrix = T) -->
<!--   } else{ -->
<!--   GSE47533 <-getGEO(filename=".data_tmp/GSE47533_series_matrix.txt.gz") -->
<!-- } -->

<!-- # Normalisation: the data is already normalised -->
<!-- expr.GSE47533 <- as.matrix(exprs(normalize.ExpressionSet.quantiles(GSE47533[[1]]))) -->
<!-- prob.GSE47533 <- unique(rownames(expr.GSE47533)) -->
<!-- data.GSE47533 <- pData(GSE47533[[1]]) -->

<!-- data.GSE47533 <- data.frame( -->
<!--                   codes = as.character(data.GSE47533$geo_accession), -->
<!--                   cell_line = "MCF7", -->
<!--                   time = data.GSE47533$`time of exposure:ch1`, -->
<!--                   condition = substr(as.character(data.GSE47533$description), 1, 4), -->
<!--                   rep = data.GSE47533$description.1) -->

<!-- data.GSE47533 <- data.GSE47533 %>% -->
<!--   mutate(rep = recode(rep, "replicate 1" = 1, -->
<!--                            "replicate 2" = 2, -->
<!--                            "replicate 3" = 3)) %>% -->
<!--   mutate_at(vars(time), as.character)  %>% -->
<!--   mutate(time = ifelse(condition == "Norm", 0, time)) -->

<!-- # Convert the probes to Symbol names -->
<!-- # The below function call will return a datafram with probe_id, gene symbol -->
<!-- # and ŕefgene_id for your data -->
<!-- anno.GSE47533 <- AnnotationDbi::select(illuminaHumanv3.db, -->
<!--        keys = prob.GSE47533, -->
<!--        columns=c("ENSEMBL", "SYMBOL", "GENENAME"), -->
<!--        keytype="PROBEID") -->

<!-- colnames(anno.GSE47533) <- c("probes", "ensgene", "symbol", "description") -->

<!-- rm(download_dir, GSE47533,  prob.GSE47533) -->

<!-- save.image("../data/data.GSE47533.Rdata") -->
<!-- ``` -->


# Load the pre-processed data

```{r}
load("../data/data.GSE47533.Rdata")
```

# Selecting the HIF Genes 
    
```{r}
# Selected genes from HIF Axis
hif.symbols <- c("TP53", "HIF1A", "EP300", "MDM2", "VHL")

hif.probes <- anno.GSE47533$probes[anno.GSE47533$symbol %in% hif.symbols]

# Select the probes and genes
expr.GSE47533.hif <- data.frame(expr.GSE47533) %>% 
  rownames_to_column('probes') %>% 
  filter(probes %in% hif.probes) %>% 
  merge(anno.GSE47533[anno.GSE47533$symbol %in% hif.symbols, c("probes","symbol")], by = "probes") %>% 
  #distinct(symbol, .keep_all = TRUE) %>% # Take the first one
  dplyr::select(!(probes)) %>% 
  arrange(symbol)
  
```

```{r}
# Function to binarize according an consensus mean of probes, add the O2 state and rename columns 

binNet <- function(b){
  
  cols <- data.GSE47533$codes %in% names(b)
  
  binarizeTimeSeries(b[,-1], method="kmeans")$binarizedMeasurements  %>% 
  as.data.frame(.)  %>% 
  aggregate(., list(symbol = b$symbol), mean) %>%  # mean of binarized probes
  mutate_at(vars(-symbol), funs(ifelse(. >= 0.5, 1, 0))) %>%  # consensus with a bies to 1 (>= 0.5)
  rbind(., c("O2", 1,0,0,0)) %>% 
    rename_at(vars(data.GSE47533$codes[cols] ),
            ~paste0(substr(data.GSE47533$condition[cols],1,2),".",
                    data.GSE47533$time[cols],".",
                    substr(data.GSE47533$cell_line[cols],1,2), ".",
                    data.GSE47533$rep[cols])) %>% 
  column_to_rownames("symbol")
  
}
```

```{r}


cols <- (data.GSE47533$rep == 1)

breast1_MCF7 <- 
expr.GSE47533.hif %>% 
  dplyr::select(c("symbol", data.GSE47533$codes[cols])) %>% arrange(symbol) %>% 
  arrange(symbol) %>% 
  rename_at(vars(data.GSE47533$codes[cols]),
            ~paste0(substr(data.GSE47533$condition[cols],1,2),".",
                    data.GSE47533$time[cols],".",
                    substr(data.GSE47533$cell_line[cols],1,2)))

knitr::kable(breast1_MCF7)

binarizeTimeSeries(breast1_MCF7[,-1], method="kmeans")$binarizedMeasurements  %>% 
  data.frame(.)  %>% 
  add_column(symbol = breast1_MCF7$symbol, .before=0) %>% 
  knitr::kable(.)
  
binarizeTimeSeries(breast1_MCF7[,-1], method="kmeans")$binarizedMeasurements  %>% 
  data.frame(.)  %>% 
  aggregate(., list(symbol = breast1_MCF7$symbol), mean) %>% 
  mutate_at(vars(-symbol), funs(ifelse(. >= 0.5, 1, 0))) %>% 
  rbind(., c("O2", 1,0,0,0)) %>% 
  knitr::kable(.)
  
```

# MCF7 breast cancer

```{r}

breast1_MCF7 <- 
expr.GSE47533.hif %>% 
  dplyr::select(c("symbol", data.GSE47533$codes[data.GSE47533$rep == 1]))  %>% 
  binNet(.) 
knitr::kable(breast1_MCF7)


breast2_MCF7 <- 
expr.GSE47533.hif %>% 
  dplyr::select(c("symbol", data.GSE47533$codes[data.GSE47533$rep == 2]))  %>% 
  binNet(.) 
knitr::kable(breast2_MCF7)


breast3_MCF7 <- 
expr.GSE47533.hif %>% 
  dplyr::select(c("symbol", data.GSE47533$codes[data.GSE47533$rep == 3]))  %>% 
  binNet(.) 
knitr::kable(breast3_MCF7)

# All breast cancer nets merged:

net <- reconstructNetwork(list(breast1_MCF7, breast2_MCF7, breast3_MCF7), method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)
plotNetworkWiring(net)
print(net)

# Individual nets of each replica:

net <- reconstructNetwork(breast1_MCF7, method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)
plotNetworkWiring(net)
print(net)

net <- reconstructNetwork(breast2_MCF7, method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)
plotNetworkWiring(net)
print(net)

net <- reconstructNetwork(breast3_MCF7, method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)
plotNetworkWiring(net)
print(net)
```


