---
title: "R Notebook"
output: html_notebook
---


```{r message=FALSE, warning=FALSE}
#setwd("~/boolnet")

if(!require("igraph")){install.packages('igraph')}

if(!require("BoolNet")){install.packages('BoolNet')}

if(!require("BiocManager")){BiocManager::install("BiocManager")}# Install bioconductor

if(!require("dplyr")){BiocManager::install("dplyr")}# Install dplyr

if(!require("Biobase")){BiocManager::install("Biobase")}# Install Biobase

if(!require("GEOquery")){BiocManager::install("GEOquery")}# Install GEOquery

```

```{r}
# Reading data
gse <- getGEO("GSE18494", destdir = "../data/GSE18494", GSEMatrix = FALSE)
gsmplatforms <- lapply(GSMList(gse),function(x) {Meta(x)$platform_id})
gsmlist = Filter(function(gsm) {Meta(gsm)$platform_id=='GPL9419'},GSMList(gse))
probesets <- Table(GPLList(gse)[[1]])$ID
data.matrix <- do.call('cbind',lapply(gsmlist,function(x) 
{tab <- Table(x)
mymatch <- match(probesets,tab$ID_REF)
return(tab$VALUE[mymatch])
}))
rownames(data.matrix) <- probesets
df <- as.data.frame(data.matrix)
```

```{r}
names(df)[names(df) == "GSM460669"] <- "norm.control.H1"
names(df)[names(df) == "GSM460670"] <- "norm.control.H2"
names(df)[names(df) == "GSM460671"] <- "norm.control.H3"
names(df)[names(df) == "GSM460672"] <- "hypo.4h.H1"
names(df)[names(df) == "GSM460673"] <- "hypo.4h.H2"
names(df)[names(df) == "GSM460674"] <- "hypo.4h.H3"
names(df)[names(df) == "GSM460675"] <- "hypo.8h.H1"
names(df)[names(df) == "GSM460676"] <- "hypo.8h.H2"
names(df)[names(df) == "GSM460677"] <- "hypo.8h.H3"
names(df)[names(df) == "GSM460678"] <- "hypo.12h.H1"
names(df)[names(df) == "GSM460679"] <- "hypo.12h.H2"
names(df)[names(df) == "GSM460680"] <- "hypo.12h.H3"


names(df)[names(df) == "GSM460681"] <- "norm.control.U1"
names(df)[names(df) == "GSM460682"] <- "norm.control.U2"
names(df)[names(df) == "GSM460683"] <- "norm.control.U3"
names(df)[names(df) == "GSM460684"] <- "hypo.4h.U1"
names(df)[names(df) == "GSM460685"] <- "hypo.4h.U2"
names(df)[names(df) == "GSM460686"] <- "hypo.4h.U3"
names(df)[names(df) == "GSM460687"] <- "hypo.8h.U1"
names(df)[names(df) == "GSM460688"] <- "hypo.8h.U2"
names(df)[names(df) == "GSM460689"] <- "hypo.8h.U3"
names(df)[names(df) == "GSM460690"] <- "hypo.12h.U1"
names(df)[names(df) == "GSM460691"] <- "hypo.12h.U2"
names(df)[names(df) == "GSM460692"] <- "hypo.12h.U3"


names(df)[names(df) == "GSM460693"] <- "norm.control.M1"
names(df)[names(df) == "GSM460694"] <- "norm.control.M2"
names(df)[names(df) == "GSM460695"] <- "norm.control.M3"
names(df)[names(df) == "GSM460696"] <- "hypo.4h.M1"
names(df)[names(df) == "GSM460697"] <- "hypo.4h.M2"
names(df)[names(df) == "GSM460698"] <- "hypo.4h.M3"
names(df)[names(df) == "GSM460699"] <- "hypo.8h.M1"
names(df)[names(df) == "GSM460700"] <- "hypo.8h.M2"
names(df)[names(df) == "GSM460701"] <- "hypo.8h.M3"
names(df)[names(df) == "GSM460702"] <- "hypo.12h.M1"
names(df)[names(df) == "GSM460703"] <- "hypo.12h.M2"
names(df)[names(df) == "GSM460704"] <- "hypo.12h.M3"
```


```{r}
# Separating replicas

# Breast Cancer
breast1 <- df[c("norm.control.M1", "hypo.4h.M1", "hypo.8h.M1","hypo.12h.M1")]
breast2 <- df[c("norm.control.M2", "hypo.4h.M2", "hypo.8h.M2","hypo.12h.M2")]
breast3 <- df[c("norm.control.M3", "hypo.4h.M3", "hypo.8h.M3","hypo.12h.M3")]


# Glioma Cancer
glioma1 <- df[c("norm.control.U1", "hypo.4h.U1", "hypo.8h.U1","hypo.12h.U1")]
glioma2 <- df[c("norm.control.U2", "hypo.4h.U2", "hypo.8h.U2","hypo.12h.U2")]
glioma3 <- df[c("norm.control.U3", "hypo.4h.U3", "hypo.8h.U3","hypo.12h.U3")]


# Hepatoma Cancer
hepatoma1 <- df[c("norm.control.H1", "hypo.4h.H1", "hypo.8h.H1","hypo.12h.H1")]
hepatoma2 <- df[c("norm.control.H2", "hypo.4h.H2", "hypo.8h.H2","hypo.12h.H2")]
hepatoma3 <- df[c("norm.control.H3", "hypo.4h.H3", "hypo.8h.H3","hypo.12h.H3")]
```


```{r}
row.names(df) = substr(row.names(df),1,nchar(row.names(df))-3)

row.names(breast1) = substr(row.names(breast1),1,nchar(row.names(breast1))-3)
row.names(breast2) = substr(row.names(breast2),1,nchar(row.names(breast2))-3)
row.names(breast3) = substr(row.names(breast3),1,nchar(row.names(breast3))-3)

row.names(glioma1) = substr(row.names(glioma1),1,nchar(row.names(glioma1))-3)
row.names(glioma2) = substr(row.names(glioma2),1,nchar(row.names(glioma2))-3)
row.names(glioma3) = substr(row.names(glioma3),1,nchar(row.names(glioma3))-3)

row.names(hepatoma1) = substr(row.names(hepatoma1),1,nchar(row.names(hepatoma1))-3)
row.names(hepatoma2) = substr(row.names(hepatoma2),1,nchar(row.names(hepatoma2))-3)
row.names(hepatoma3) = substr(row.names(hepatoma3),1,nchar(row.names(hepatoma3))-3)

mart<- biomaRt::useMart(biomart = 'ensembl', dataset = 'hsapiens_gene_ensembl')

# mapping example
refseq_mapping <- biomaRt::getBM(attributes = c("refseq_mrna","hgnc_symbol"), 
                        filters="refseq_mrna", # you swap out of this filter for whatever your input is
                        values=row.names(df), # vector of your NMf
                        mart=mart)

refseq_mapping <- as.data.frame(refseq_mapping)
names(refseq_mapping)[names(refseq_mapping) == "refseq_mrna"] <- "ID_REF"
df$ID_REF <- row.names(df)
df <- merge(df, refseq_mapping, by = "ID_REF")

breast1$ID_REF <- row.names(breast1)
breast1 <- merge(breast1, refseq_mapping, by = "ID_REF")
breast2$ID_REF <- row.names(breast2)
breast2 <- merge(breast2, refseq_mapping, by = "ID_REF")
breast3$ID_REF <- row.names(breast3)
breast3 <- merge(breast3, refseq_mapping, by = "ID_REF")

glioma1$ID_REF <- row.names(glioma1)
glioma1 <- merge(glioma1, refseq_mapping, by = "ID_REF")
glioma2$ID_REF <- row.names(glioma2)
glioma2 <- merge(glioma2, refseq_mapping, by = "ID_REF")
glioma3$ID_REF <- row.names(glioma3)
glioma3 <- merge(glioma3, refseq_mapping, by = "ID_REF")

hepatoma1$ID_REF <- row.names(hepatoma1)
hepatoma1 <- merge(hepatoma1, refseq_mapping, by = "ID_REF")
hepatoma2$ID_REF <- row.names(hepatoma2)
hepatoma2 <- merge(hepatoma2, refseq_mapping, by = "ID_REF")
hepatoma3$ID_REF <- row.names(hepatoma3)
hepatoma3 <- merge(hepatoma3, refseq_mapping, by = "ID_REF")
```

```{r}
saveRDS(df, file = "../data/GSE18494.rds")
#df <- readRDS("GSE47534.rds")
```

```{r}
breast1 <- breast1[ which(breast1$hgnc_symbol=='HIF1A'
                      | breast1$hgnc_symbol=='EP300'
                      | breast1$hgnc_symbol=='VHL'
                      | breast1$hgnc_symbol=='MDM2'
                      | breast1$hgnc_symbol=='TP53'), ]
breast1 <- unique( breast1[ , 2:ncol(breast1)] )

breast2 <- breast2[ which(breast2$hgnc_symbol=='HIF1A'
                      | breast2$hgnc_symbol=='EP300'
                      | breast2$hgnc_symbol=='VHL'
                      | breast2$hgnc_symbol=='MDM2'
                      | breast2$hgnc_symbol=='TP53'), ]
breast2 <- unique( breast2[ , 2:ncol(breast2)] )

breast3 <- breast3[ which(breast3$hgnc_symbol=='HIF1A'
                      | breast3$hgnc_symbol=='EP300'
                      | breast3$hgnc_symbol=='VHL'
                      | breast3$hgnc_symbol=='MDM2'
                      | breast3$hgnc_symbol=='TP53'), ]
breast3 <- unique( breast3[ , 2:ncol(breast3)] )

glioma1 <- glioma1[ which(glioma1$hgnc_symbol=='HIF1A'
                      | glioma1$hgnc_symbol=='EP300'
                      | glioma1$hgnc_symbol=='VHL'
                      | glioma1$hgnc_symbol=='MDM2'
                      | glioma1$hgnc_symbol=='TP53'), ]
glioma1 <- unique( glioma1[ , 2:ncol(glioma1)] )

glioma2 <- glioma2[ which(glioma2$hgnc_symbol=='HIF1A'
                      | glioma2$hgnc_symbol=='EP300'
                      | glioma2$hgnc_symbol=='VHL'
                      | glioma2$hgnc_symbol=='MDM2'
                      | glioma2$hgnc_symbol=='TP53'), ]
glioma2 <- unique( glioma2[ , 2:ncol(glioma2)] )

glioma3 <- glioma3[ which(glioma3$hgnc_symbol=='HIF1A'
                      | glioma3$hgnc_symbol=='EP300'
                      | glioma3$hgnc_symbol=='VHL'
                      | glioma3$hgnc_symbol=='MDM2'
                      | glioma3$hgnc_symbol=='TP53'), ]
glioma3 <- unique( glioma3[ , 2:ncol(glioma3)] )


hepatoma1 <- hepatoma1[ which(hepatoma1$hgnc_symbol=='HIF1A'
                      | hepatoma1$hgnc_symbol=='EP300'
                      | hepatoma1$hgnc_symbol=='VHL'
                      | hepatoma1$hgnc_symbol=='MDM2'
                      | hepatoma1$hgnc_symbol=='TP53'), ]
hepatoma1 <- unique( hepatoma1[ , 2:ncol(hepatoma1)] )

hepatoma2 <- hepatoma2[ which(hepatoma2$hgnc_symbol=='HIF1A'
                      | hepatoma2$hgnc_symbol=='EP300'
                      | hepatoma2$hgnc_symbol=='VHL'
                      | hepatoma2$hgnc_symbol=='MDM2'
                      | hepatoma2$hgnc_symbol=='TP53'), ]
hepatoma2 <- unique( hepatoma2[ , 2:ncol(hepatoma2)] )

hepatoma3 <- hepatoma3[ which(hepatoma3$hgnc_symbol=='HIF1A'
                      | hepatoma3$hgnc_symbol=='EP300'
                      | hepatoma3$hgnc_symbol=='VHL'
                      | hepatoma3$hgnc_symbol=='MDM2'
                      | hepatoma3$hgnc_symbol=='TP53'), ]
hepatoma3 <- unique( hepatoma3[ , 2:ncol(hepatoma3)] )
```

1559227_s_at	ENSG00000134086	VHL	von Hippel-Lindau tumor suppressor
	
201472_at	ENSG00000155959	VBP1	VHL binding protein 1

203844_at	ENSG00000134086	VHL	von Hippel-Lindau tumor suppressor

Não encontrei o VHL nesse dataset
Possiveis aliases segundo gene cards -> PVHL, Von Hippel-Lindau Syndrome, Elongin Binding Protein, HRCA1, RCA1, VHLL

HIF1A duplicado

```{r}
row.names(breast1) <- breast1$hgnc_symbol
breast1 <- breast1[, -5]  
bin <- binarizeTimeSeries(breast1, method="kmeans")$binarizedMeasurements
bin = rbind(bin,c(1,0,0,0))
row.names(bin)[5] <- "O2"
net <- reconstructNetwork(bin, method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)
plotNetworkWiring(net)
print(net)
```

```{r}
row.names(breast2) <- breast2$hgnc_symbol
breast2 <- breast2[, -5]  
bin <- binarizeTimeSeries(breast2, method="kmeans")$binarizedMeasurements
bin = rbind(bin,c(1,0,0,0))
row.names(bin)[5] <- "O2"
net <- reconstructNetwork(bin, method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)
plotNetworkWiring(net)
print(net)
```

```{r}
row.names(breast3) <- breast3$hgnc_symbol
breast3 <- breast3[, -5]  
bin <- binarizeTimeSeries(breast3, method="kmeans")$binarizedMeasurements
bin = rbind(bin,c(1,0,0,0))
row.names(bin)[5] <- "O2"
net <- reconstructNetwork(bin, method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)
plotNetworkWiring(net)
print(net)
```

```{r}
row.names(glioma1) <- glioma1$hgnc_symbol
glioma1 <- glioma1[, -5]  
bin <- binarizeTimeSeries(glioma1, method="kmeans")$binarizedMeasurements
bin = rbind(bin,c(1,0,0,0))
row.names(bin)[5] <- "O2"
net <- reconstructNetwork(bin, method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)
plotNetworkWiring(net)
print(net)
```

```{r}
row.names(glioma2) <- glioma2$hgnc_symbol
glioma2 <- glioma2[, -5]  
bin <- binarizeTimeSeries(glioma2, method="kmeans")$binarizedMeasurements
bin = rbind(bin,c(1,0,0,0))
row.names(bin)[5] <- "O2"
net <- reconstructNetwork(bin, method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)
plotNetworkWiring(net)
print(net)
```

```{r}
row.names(glioma3) <- glioma3$hgnc_symbol
glioma3 <- glioma3[, -5]  
bin <- binarizeTimeSeries(glioma3, method="kmeans")$binarizedMeasurements
bin = rbind(bin,c(1,0,0,0))
row.names(bin)[5] <- "O2"
net <- reconstructNetwork(bin, method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)
plotNetworkWiring(net)
print(net)
```

```{r}
row.names(hepatoma1) <- hepatoma1$hgnc_symbol
hepatoma1 <- hepatoma1[, -5]  
bin <- binarizeTimeSeries(hepatoma1, method="kmeans")$binarizedMeasurements
bin = rbind(bin,c(1,0,0,0))
row.names(bin)[5] <- "O2"
net <- reconstructNetwork(bin, method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)
plotNetworkWiring(net)
print(net)
```

```{r}
row.names(hepatoma2) <- hepatoma2$hgnc_symbol
hepatoma2 <- hepatoma2[, -5]  
bin <- binarizeTimeSeries(hepatoma2, method="kmeans")$binarizedMeasurements
bin = rbind(bin,c(1,0,0,0))
row.names(bin)[5] <- "O2"
net <- reconstructNetwork(bin, method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)
plotNetworkWiring(net)
print(net)
```

```{r}
row.names(hepatoma3) <- hepatoma3$hgnc_symbol
hepatoma3 <- hepatoma3[, -5]  
bin <- binarizeTimeSeries(hepatoma3, method="kmeans")$binarizedMeasurements
bin = rbind(bin,c(1,0,0,0))
row.names(bin)[5] <- "O2"
net <- reconstructNetwork(bin, method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)
plotNetworkWiring(net)
print(net)
```

```{r}
# Reading data
gse <- getGEO("GSE47534", destdir = "..", GSEMatrix = FALSE)
gsmplatforms <- lapply(GSMList(gse),function(x) {Meta(x)$platform_id})
gsmlist = Filter(function(gsm) {Meta(gsm)$platform_id=='GPL6884'},GSMList(gse))
probesets <- Table(GPLList(gse)[[1]])$ID
data.matrix <- do.call('cbind',lapply(gsmlist,function(x) 
{tab <- Table(x)
mymatch <- match(probesets,tab$ID_REF)
return(tab$VALUE[mymatch])
}))
rownames(data.matrix) <- probesets
df <- as.data.frame(data.matrix)
```

```{r}

```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
