---
title: "Merging Breast Cancer Cell-lines (E-GEOD-18494, GSE47533 and GSE41491)"
output:
  pdf_document: default
  github_document: 
    df_print: paged
    html_preview: FALSE
    keep_html: FALSE
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_format = "all") })    
---
```{r error=TRUE, message=FALSE, warning=FALSE, include=FALSE, purl=FALSE, results='hide'}
## This chunk automatically generates a text .R version of this script when running within knitr.
# input  = knitr::current_input()  # filename of input document
# output = paste(tools::file_path_sans_ext(input), 'R', sep = '.')
# try(knitr::purl(input,output,documentation=2,quiet=T), silent = TRUE)
## Avoid duplicate label error of knitr::purl
options(knitr.duplicate.label = 'allow')
## Code to browse the markdown file with renderized images.
knitr::opts_chunk$set(
  fig.path = "figs/BoolNetInfer-"
)
```


```{r message=FALSE, warning=FALSE, include=FALSE}

#packages_cran = c("igraph", "BoolNet", "BiocManager", "tidyverse", "fs", "ff","RSQLite")
packages_cran = c("igraph", "BoolNet", "BiocManager", "tidyverse", "ggraph")

# Install and load packages
package.check <- lapply(packages_cran, FUN = function(x) {
  if (!require(x, character.only = TRUE)) {
    install.packages(x, dependencies = TRUE)
    library(x, character.only = TRUE)
  }
})

if(!require('cowplot')) {install.packages('cowplot')} # provides various features that help with creating publication-quality figures
if(!require('grid')){install.packages('grid')} # 
if(!require('gridExtra')) {install.packages('gridExtra')} #  Provides a number of user-level functions to work with "grid" graphics, notably to arrange multiple grid-based plots on a page, and draw tables.
if(!require('devtools')) {install.packages("devtools", dependencies = TRUE)}
if(!require('TDM')) {devtools::install_github("greenelab/TDM")}


# For oligo and ArrayExpress First install:
#install.packages('https://cran.r-project.org/src/contrib/Archive/ff/ff_2.2-14.tar.gz',repos=NULL)
# 
# if(!require("hgu133plus2hsentrezg.db")){ 
#   install.packages('http://mbni.org/customcdf/13.0.0/entrezg.download/hgu133plus2hsentrezg.db_13.0.0.tar.gz',repos=NULL) 
# }
# 
#packages_bioconductor = c("Biobase", "GEOquery", "affyPLM", "ArrayExpress", "illuminaHumanv3.db", "hgu133plus2.db")
packages_bioconductor = c()

# Install and load packages
package.check <- lapply(packages_bioconductor, FUN = function(x) {
  if (!require(x, character.only = TRUE)) {
    BiocManager::install(x, dependencies = TRUE)
    library(x, character.only = TRUE)
  }
})

rm(package.check, packages_bioconductor, packages_cran)

```


```{r include=FALSE}
load("../data/data.GSE47533.Rdata")
load("../data/data.EGEOD18494.Rdata")
load("../data/data.GSE41491.Rdata")
```


Rede Alexandre (reunião 27-11-2020)

HIF1a, !VHL & !O2  
p53, !Mdm2   
Mdm2, p53 & !VHL  
VHL, HIF1a & !p53   
p300, ((p53 & HIF1a) & !VHL) | (!(p53 & HIF1a) & VHL)  
BIM, !MCL_1 & !BCLXL & !BCL2  
BAD, p53  
BID, (!HIF1a & (p53 & VHL)) | (!MCL_1 & !BCLXL & !BCL2)  
BIK, !MCL_1 & !BCLXL & !BCL2  
MCL_1, HIF1a   
BCLXL, HIF1a & !(p53 & VHL) & ((!Casp3 & !BAD) | (!Casp3 & BCL2))  
BCL2, HIF1a & !(p53 & VHL) & ((MCL_1 & !BIM & !BIK & !BAD) | (!BIM & !BIK & BCLXL & !BAD))  
IAPs, !DIABLO  
BAX, (BIM & !BCLXL) | (BIK & !BCLXL & !BCL2) | (BID & !BCLXL & !BCL2) | (BIM & BID) | (BIM & BIK) | (BIM & !BCL2) | (!MCL_1 & BIM)  
BAK, (!MCL_1 & BIM & !BCLXL) | (BID & !BCL2) | (BID & !BCLXL) | (!MCL_1 & BID) | (!MCL_1 & BIK & !BCLXL) | (BIM & BID) | (BIK & BID)  
DIABLO, BAX | BAK  
Cyto_C, BAX | BAK  
Casp9, Casp3 | (!IAPs & Cyto_C)  
Casp3, !IAPs & Casp9  

BCLXL ?

No VHL in GSE41491

```{r}
# Selected genes from HIF Axis
hif.symbols <- c("HIF1A", "TP53", "MDM2", "VHL", "EP300","TMBIM1","TMBIM4", "TMBIM6","BAD","BIK","MCL1", "BCL2","BIRC2", "BIRC3", "BIRC5", "BIRC6", "BIRC7", "DIABLO", "BAX", "CYC1", "CASP9", "CASP3") 

hif.probes <- anno.EGEOD18494$probes[anno.EGEOD18494$symbol %in% hif.symbols]

# Select the probes and genes
# EGEOD18494
expr.EGEOD18494.hif <- as.data.frame(expr.EGEOD18494) %>% 
  rownames_to_column('probes') %>% 
  filter(probes %in% hif.probes) %>% 
  merge(anno.EGEOD18494[anno.EGEOD18494$symbol %in% hif.symbols, c("probes","symbol")], by = "probes") %>% 
  mutate(., symbol=ifelse(symbol %in% c("TMBIM1","TMBIM4", "TMBIM6"), "BIM", symbol)) %>%
  mutate(., symbol=ifelse(symbol %in% c("BIRC2", "BIRC3", "BIRC5", "BIRC6", "BIRC7"), "IAPs", symbol)) %>%
  group_by(symbol) %>%
  summarise_at(vars(-probes), funs(mean(., na.rm=TRUE)))  %>%
  column_to_rownames(var = "symbol") %>%
  dplyr::select(c(data.EGEOD18494$codes[data.EGEOD18494$cell_line == "MDA-MB231 breast cancer"]))


hif.probes <- anno.GSE47533$probes[anno.GSE47533$symbol %in% hif.symbols]

# GSE47533
expr.GSE47533.hif <- as.data.frame(expr.GSE47533) %>% 
  rownames_to_column('probes') %>% 
  filter(probes %in% hif.probes) %>% 
  merge(anno.GSE47533[anno.GSE47533$symbol %in% hif.symbols, c("probes","symbol")], by = "probes") %>% 
  mutate(., symbol=ifelse(symbol %in% c("TMBIM1","TMBIM4", "TMBIM6"), "BIM", symbol)) %>%
  mutate(., symbol=ifelse(symbol %in% c("BIRC2", "BIRC3", "BIRC5", "BIRC6", "BIRC7"), "IAPs", symbol)) %>%
  group_by(symbol) %>%
  summarise_at(vars(-probes), funs(mean(., na.rm=TRUE)))  %>%
  column_to_rownames(var = "symbol") 
  

hif.probes <- anno.GSE41491$probes[anno.GSE41491$symbol %in% hif.symbols]

# GSE41491
expr.GSE41491.hif <- as.data.frame(expr.GSE41491) %>% 
  rownames_to_column('probes') %>% 
  filter(probes %in% hif.probes) %>% 
  merge(anno.GSE41491[anno.GSE41491$symbol %in% hif.symbols, c("probes","symbol")], by = "probes") %>% 
  mutate(., symbol=ifelse(symbol %in% c("TMBIM1","TMBIM4", "TMBIM6"), "BIM", symbol)) %>%
  mutate(., symbol=ifelse(symbol %in% c("BIRC2", "BIRC3", "BIRC5", "BIRC6", "BIRC7"), "IAPs", symbol)) %>%
  group_by(symbol) %>%
  summarise_at(vars(-probes), funs(mean(., na.rm=TRUE)))  %>%
  column_to_rownames(var = "symbol") 

```


```{r}

write.table(expr.GSE47533.hif,"expr.GSE47533.txt", sep ="\t")
write.table(expr.EGEOD18494.hif, "expr.EGEOD18494.txt", sep ="\t")
write.table(expr.GSE41491.hif, "expr.GSE41491.txt", sep ="\t")

expr.EGEOD18494.tdm <- tdm_transform(ref_file = "expr.GSE47533.txt", file = "expr.EGEOD18494.txt")

expr.GSE41491.tdm <- tdm_transform(ref_file = "expr.GSE47533.txt", file = "expr.GSE41491.txt")

symbols <-  expr.EGEOD18494.tdm$gene
expr.EGEOD18494.tdm$gene  <- NULL

expr.EGEOD18494.tdm <- as.data.frame(matrix(as.numeric(unlist(expr.EGEOD18494.tdm)),  
                                            nrow = dim(expr.EGEOD18494.tdm)[1], 
                                            ncol = dim(expr.EGEOD18494.tdm)[2]))
colnames(expr.EGEOD18494.tdm) <- colnames(expr.EGEOD18494.hif)

rownames(expr.EGEOD18494.tdm) <- symbols
```

```{r}
row_medians_assayData <- 
  Biobase::rowMedians(as.matrix(expr.GSE47533))

RLE_data <- sweep(expr.GSE47533, 1, row_medians_assayData)

RLE_data <- as.data.frame(RLE_data)
RLE_data_gathered <- 
  tidyr::gather(RLE_data, patient_array, log2_expression_deviation)

ggplot2::ggplot(RLE_data_gathered, aes(patient_array,
                                       log2_expression_deviation)) + 
  geom_boxplot(outlier.shape = NA) + 
  ylim(c(-2, 2)) + 
  ggtitle("expr.GSE47533") + 
  theme(axis.text.x = element_text(colour = "aquamarine4", 
                                  angle = 60, size = 6.5, hjust = 1 ,
                                  face = "bold"))


row_medians_assayData <- 
  Biobase::rowMedians(as.matrix(expr.EGEOD18494.tdm))

RLE_data <- sweep(expr.EGEOD18494.tdm, 1, row_medians_assayData)

RLE_data <- as.data.frame(RLE_data)
RLE_data_gathered <- 
  tidyr::gather(RLE_data, patient_array, log2_expression_deviation)

ggplot2::ggplot(RLE_data_gathered, aes(patient_array,
                                       log2_expression_deviation)) + 
  geom_boxplot(outlier.shape = NA) + 
  ylim(c(-2, 2)) + 
  ggtitle("expr.EGEOD18494.tdm") + 
  theme(axis.text.x = element_text(colour = "aquamarine4", 
                                  angle = 60, size = 6.5, hjust = 1 ,
                                  face = "bold"))

row_medians_assayData <- 
  Biobase::rowMedians(as.matrix(expr.EGEOD18494.hif))

RLE_data <- sweep(expr.EGEOD18494.hif, 1, row_medians_assayData)

RLE_data <- as.data.frame(RLE_data)
RLE_data_gathered <- 
  tidyr::gather(RLE_data, patient_array, log2_expression_deviation)

ggplot2::ggplot(RLE_data_gathered, aes(patient_array,
                                       log2_expression_deviation)) + 
  geom_boxplot(outlier.shape = NA) + 
  ylim(c(-2, 2)) + 
  ggtitle("expr.EGEOD18494.hif") + 
  theme(axis.text.x = element_text(colour = "aquamarine4", 
                                  angle = 60, size = 6.5, hjust = 1 ,
                                  face = "bold"))


rm(RLE_data, RLE_data_gathered, row_medians_assayData)
```

```{r}
hif.probes <- anno.GSE47533$probes[anno.GSE47533$symbol %in% hif.symbols]


row_medians_assayData <- 
  Biobase::rowMedians(as.matrix(t(expr.GSE47533.hif)))

RLE_data <- sweep(t(expr.GSE47533.hif), 1, row_medians_assayData)

RLE_data <- as.data.frame(RLE_data)
RLE_data_gathered <- 
  tidyr::gather(RLE_data, patient_array, log2_expression_deviation)

ggplot2::ggplot(RLE_data_gathered, aes(patient_array,
                                       log2_expression_deviation)) + 
  geom_boxplot(outlier.shape = NA) + 
  ylim(c(-4, 5)) + 
  ggtitle("GSE47533 MCF7 - Raw data") +
  theme(axis.text.x = element_text(colour = "aquamarine4", 
                                  angle = 60, size = 6.5, hjust = 1 ,
                                  face = "bold"))


row_medians_assayData <- 
  Biobase::rowMedians(as.matrix(t(expr.EGEOD18494.hif)))

RLE_data <- sweep(t(expr.EGEOD18494.hif), 1, row_medians_assayData)

RLE_data <- as.data.frame(RLE_data)
RLE_data_gathered <- 
  tidyr::gather(RLE_data, patient_array, log2_expression_deviation)

ggplot2::ggplot(RLE_data_gathered, aes(patient_array,
                                       log2_expression_deviation)) + 
  geom_boxplot(outlier.shape = NA) + 
  ggtitle("EGEOD18494 MDA-MB231 - Raw data") +
  ylim(c(-4, 5)) + 
  theme(axis.text.x = element_text(colour = "aquamarine4", 
                                  angle = 60, size = 6.5, hjust = 1 ,
                                  face = "bold"))

row_medians_assayData <- 
  Biobase::rowMedians(as.matrix(t(expr.EGEOD18494.tdm)))

RLE_data <- sweep(t(expr.EGEOD18494.tdm), 1, row_medians_assayData)

RLE_data <- as.data.frame(RLE_data)
RLE_data_gathered <- 
  tidyr::gather(RLE_data, patient_array, log2_expression_deviation)

ggplot2::ggplot(RLE_data_gathered, aes(patient_array,
                                       log2_expression_deviation)) + 
  geom_boxplot(outlier.shape = NA) + 
  ylim(c(-4, 5)) + 
  ggtitle("EGEOD18494 MDA-MB231 - TDM data") +
  theme(axis.text.x = element_text(colour = "aquamarine4", 
                                  angle = 60, size = 6.5, hjust = 1 ,
                                  face = "bold"))

rm(RLE_data, RLE_data_gathered, row_medians_assayData)
```


```{r}
require(BiTrinA)
```
# 

```{r}

expr.GSE47533.hif.bin <- binarizeMatrix(expr.GSE47533.hif, 
               method = c("BASCA"), 
               adjustment = "none")

expr.GSE47533.hif.bin$symbol <- row.names(expr.GSE47533.hif.bin)

expr.GSE47533.hif.bin <- expr.GSE47533.hif.bin[, c(as.character(data.GSE47533$codes), c("threshold", "p.value", "symbol"))]

names(expr.GSE47533.hif.bin) <- c(paste0(substr(data.GSE47533$condition,1,4),".", data.GSE47533$time, ".", data.GSE47533$rep), c("threshold", "p.value", "symbol"))

head(expr.GSE47533.hif.bin) %>% 
  knitr::kable(.)
```


```{r}
expr.GSE47533.hif.mean <- expr.GSE47533.hif.bin %>%
 mutate(Norm = rowMeans(dplyr::select(., starts_with("Norm"))),
        Hypo.16h = rowMeans(dplyr::select(., starts_with("Hypo.16h"))),
        Hypo.32h = rowMeans(dplyr::select(., starts_with("Hypo.32h"))),
        Hypo.48h = rowMeans(dplyr::select(., starts_with("Hypo.48h"))))  %>%
  dplyr::select(., -ends_with(c(".1",".2", ".3")))



expr.GSE47533.hif.pivot <- expr.GSE47533.hif.mean %>%
  group_by(symbol) %>%
  pivot_longer(cols = starts_with(c("Norm","Hypo")), names_to = "codes", values_to = "value")

expr.GSE47533.hif.pivot$codes <- factor(expr.GSE47533.hif.pivot$codes,  levels =  c("Norm", "Hypo.16h" , "Hypo.32h" , "Hypo.48h"))

expr.GSE47533.hif.pivot$time <- as.numeric(expr.GSE47533.hif.pivot$codes)

# hif.symbols <- c("HIF1A", "TP53", "MDM2", "VHL", "EP300","TMBIM1","TMBIM4", "TMBIM6","BAD","BIK","MCL1", "BCL2","BIRC2", "BIRC3", "BIRC5", "BIRC6", "BIRC7", "DIABLO", "BAX", "CYC1", "CASP9", "CASP3") 

p.MCF7 <- ggplot(aes(x = factor(time), y = value, group = symbol, color="red"),  
           data = expr.GSE47533.hif.pivot[expr.GSE47533.hif.pivot$symbol %in% c("HIF1A", "TP53", "MDM2", "VHL", "EP300","BIM","BAD","BIK","MCL1", "BCL2","IAPs", "DIABLO", "BAX", "CYC1", "CASP9", "CASP3"),]) +
  geom_point() + 
  geom_line() + 
  scale_x_discrete(breaks = c(1, 2, 3, 4), 
                 labels = c("Normoxia", "Hypoxia: 16h" , "Hypoxia: 32h" , "Hypoxia: 48h")) +
  xlab("Conditions") + ylab("Gene Expression") + 
  ggtitle("GSE47533 MCF7 - Raw data") +
  theme(legend.position = "none", axis.text.x=element_text(color = "black", size=7, angle=30, vjust=.8, hjust=0.8)) +
  #geom_line(aes(linetype=Symbol, color=Symbol)) +
  facet_wrap(~ symbol) 

p.MCF7
```

# EGEOD18494


```{r}

# expr.EGEOD18494.hif.bin <- binarizeMatrix(expr.EGEOD18494.tdm, 
#                method = c("BASCA"),
#                tau = 0.15,
#                #sigma = 0.9,
#                adjustment = "none")

expr.EGEOD18494.hif.bin <- binarizeMatrix(expr.EGEOD18494.hif)

#expr.EGEOD18494.hif.bin <- expr.EGEOD18494.tdm

expr.EGEOD18494.hif.bin$symbol <- row.names(expr.EGEOD18494.hif.bin)

row <- data.EGEOD18494$cell_line == "MDA-MB231 breast cancer"
expr.EGEOD18494.hif.bin <- expr.EGEOD18494.hif.bin[, c(as.character(data.EGEOD18494$codes[row]), c("threshold", "p.value", "symbol"))]

names(expr.EGEOD18494.hif.bin) <- c(paste0(substr(data.EGEOD18494$condition[row],1,4),".", data.EGEOD18494$time[row], ".", data.EGEOD18494$rep[row]), c("threshold", "p.value", "symbol"))

#names(expr.EGEOD18494.hif.bin) <- c(paste0(substr(data.EGEOD18494$condition[row],1,4),".", data.EGEOD18494$time[row], ".", data.EGEOD18494$rep[row]), c("symbol"))

# head(expr.EGEOD18494.hif.bin) %>% 
#   knitr::kable(.)


expr.EGEOD18494.hif.mean <- expr.EGEOD18494.hif.bin %>%
 mutate(norm = rowMeans(dplyr::select(., starts_with("norm"))),
        hypo.4h = rowMeans(dplyr::select(., starts_with("hypo.4h"))),
        hypo.8h = rowMeans(dplyr::select(., starts_with("hypo.8h"))),
        hypo.12h = rowMeans(dplyr::select(., starts_with("hypo.12h"))))  %>%
  dplyr::select(., -ends_with(c(".1",".2", ".3")))



expr.EGEOD18494.hif.pivot <- expr.EGEOD18494.hif.mean %>%
  group_by(symbol) %>%
  pivot_longer(cols = starts_with(c("Norm","Hypo")), names_to = "codes", values_to = "value")

expr.EGEOD18494.hif.pivot$codes <- factor(expr.EGEOD18494.hif.pivot$codes,  levels =  c("norm", "hypo.4h" , "hypo.8h" , "hypo.12h"))

expr.EGEOD18494.hif.pivot$time <- as.numeric(expr.EGEOD18494.hif.pivot$codes)

# hif.symbols <- c("HIF1A", "TP53", "MDM2", "VHL", "EP300","TMBIM1","TMBIM4", "TMBIM6","BAD","BIK","MCL1", "BCL2","BIRC2", "BIRC3", "BIRC5", "BIRC6", "BIRC7", "DIABLO", "BAX", "CYC1", "CASP9", "CASP3") 

p.MDA <- ggplot(aes(x = factor(time), y = value, group = symbol, color="red"),  
           data = expr.EGEOD18494.hif.pivot[expr.EGEOD18494.hif.pivot$symbol %in% c("HIF1A", "TP53", "MDM2", "VHL", "EP300","BIM","BAD","BIK","MCL1", "BCL2","IAPs", "DIABLO", "BAX", "CYC1", "CASP9", "CASP3"),]) +
  geom_point() + 
  geom_line() + 
  scale_x_discrete(breaks = c(1, 2, 3, 4), 
                 labels = c("Normoxia", "Hypoxia: 4h" , "Hypoxia: 8h" , "Hypoxia: 12h")) +
  xlab("Conditions") + ylab("Gene Expression") +
  ggtitle("EGEOD18494 MDA-MB231 - Raw data") +
  theme(legend.position = "none", axis.text.x=element_text(color = "black", size=7, angle=30, vjust=.8, hjust=0.8)) +
  #geom_line(aes(linetype=Symbol, color=Symbol)) +
  facet_wrap(~ symbol) 

p.MDA

par(mfrow = c(1,2))

library(cowplot)

plot_grid(p.MDA, p.MCF7, labels = c('A', 'B'))

```


```{r}

library("pheatmap")
library("ComplexHeatmap")

data.EGEOD18494$time <- factor(data.EGEOD18494$time,  levels =  c("control", "4h" , "8h" , "12h"))

row <- data.EGEOD18494$cell_line == "MDA-MB231 breast cancer"

annotation_for_heatmap <- droplevels(data.frame(time = data.EGEOD18494$time[row], condition = data.EGEOD18494$condition[row]))

#row.names(annotation_for_heatmap) <- colnames(expr.EGEOD18494.hif)
row.names(annotation_for_heatmap) <- paste0(substr(data.EGEOD18494$condition[row],1,4),".", data.EGEOD18494$time[row], ".", data.EGEOD18494$rep[row])

dists <- as.matrix(dist(t(expr.EGEOD18494.hif), method = "manhattan")) 

#rownames(dists) <- colnames(expr.EGEOD18494.hif)

rownames(dists) <- c(paste0(substr(data.EGEOD18494$condition[row],1,4),".", data.EGEOD18494$time[row], ".", data.EGEOD18494$rep[row]))
colnames(dists) <- c(paste0(substr(data.EGEOD18494$condition[row],1,4),".", data.EGEOD18494$time[row], ".", data.EGEOD18494$rep[row]))

hmcol <- rev(colorRampPalette(RColorBrewer::brewer.pal(9, "YlOrRd"))(255))

diag(dists) <- NA 

ann_colors <- list(  
  time = RColorBrewer::brewer.pal(length(levels(data.EGEOD18494$time)), "Set2"),
  condition = c("red", "blue")
)

ann_colors
names(ann_colors$time) <- levels(data.EGEOD18494$time)
names(ann_colors$condition) <- levels(data.EGEOD18494$condition)

pheatmap(dists, col = (hmcol), 
         annotation_row = annotation_for_heatmap,
         annotation_colors = ann_colors,
         legend = TRUE, 
         treeheight_row = 0,
         legend_breaks = c(min(dists, na.rm = TRUE), 
                           max(dists, na.rm = TRUE)), 
         legend_labels = (c("small dist", "large dist")),
         main = "Clustering of Samples (EGEOD18494)")

```

```{r}

row <- data.EGEOD18494$cell_line == "MDA-MB231 breast cancer"

annotation_for_heatmap <- droplevels(data.frame(time = data.EGEOD18494$time[row], condition = data.EGEOD18494$condition[row]))

row.names(annotation_for_heatmap) <- colnames(expr.EGEOD18494.hif)

dists <- as.matrix(dist(expr.EGEOD18494.hif, method = "euclidean")) 

rownames(dists) <- rownames(expr.EGEOD18494.hif)

hmcol <- rev(colorRampPalette(RColorBrewer::brewer.pal(9, "YlOrRd"))(255))
colnames(dists) <- rownames(expr.EGEOD18494.hif)
diag(dists) <- NA 

ann_colors <- list(  
  time = RColorBrewer::brewer.pal(length(levels(data.EGEOD18494$time)), "Set2"),
  condition = c("#EF8A62", "#67A9CF")
)

ann_colors
names(ann_colors$time) <- levels(data.EGEOD18494$time)
names(ann_colors$condition) <- levels(data.EGEOD18494$condition)

pheatmap(dists, #row = (hmcol), 
         #annotation_col = annotation_for_heatmap,
         #annotation_colors = ann_colors,
         legend = TRUE, 
         treeheight_row = 0,
         legend_breaks = c(min(dists, na.rm = TRUE), 
                           max(dists, na.rm = TRUE)), 
         legend_labels = (c("small dist", "large dist")),
         main = "Clustering of Gene Expression \n Euclidian Distance  (EGEOD18494)")

#---------------------------------------------------------------------------------

row <- data.EGEOD18494$cell_line == "MDA-MB231 breast cancer" &  data.EGEOD18494$condition == "hypoxia"

annotation_for_heatmap <- droplevels(data.frame(time = data.EGEOD18494$time[row], condition = data.EGEOD18494$condition[row]))

expr.row <- (colnames(expr.EGEOD18494.hif) %in% data.EGEOD18494$codes[data.EGEOD18494$cell_line == "MDA-MB231 breast cancer" &  data.EGEOD18494$condition == "hypoxia"])

row.names(annotation_for_heatmap) <- colnames(expr.EGEOD18494.hif[expr.row])

dists <- as.matrix(dist(expr.EGEOD18494.hif[expr.row], method = "euclidean")) 

rownames(dists) <- rownames(expr.EGEOD18494.hif[expr.row])

hmcol <- rev(colorRampPalette(RColorBrewer::brewer.pal(9, "YlOrRd"))(255))
colnames(dists) <- rownames(expr.EGEOD18494.hif[expr.row])
diag(dists) <- NA 

ann_colors <- list(  
  time = RColorBrewer::brewer.pal(length(levels(data.EGEOD18494$time)), "Set2"),
  condition = c("#EF8A62", "#67A9CF")
)

ann_colors
names(ann_colors$time) <- levels(data.EGEOD18494$time)
names(ann_colors$condition) <- levels(data.EGEOD18494$condition)

p1 <- pheatmap(dists, #row = (hmcol), 
         #annotation_col = annotation_for_heatmap,
         #annotation_colors = ann_colors,
         legend = TRUE, 
         treeheight_row = 0,
         legend_breaks = c(min(dists, na.rm = TRUE), 
                           max(dists, na.rm = TRUE)), 
         legend_labels = (c("small dist", "large dist")),
         main = "Clustering of Gene Expression on Hypoxia \n  Euclidian Distance  (EGEOD18494)")

#---------------------------------------------------------------------------------

row <- data.EGEOD18494$cell_line == "MDA-MB231 breast cancer" &  data.EGEOD18494$condition == "normoxia"

annotation_for_heatmap <- droplevels(data.frame(time = data.EGEOD18494$time[row], condition = data.EGEOD18494$condition[row]))

expr.row <- (colnames(expr.EGEOD18494.hif) %in% data.EGEOD18494$codes[data.EGEOD18494$cell_line == "MDA-MB231 breast cancer" &  data.EGEOD18494$condition == "normoxia"])

row.names(annotation_for_heatmap) <- colnames(expr.EGEOD18494.hif[expr.row])

dists <- as.matrix(dist(expr.EGEOD18494.hif[expr.row], method = "euclidean")) 

rownames(dists) <- rownames(expr.EGEOD18494.hif[expr.row])

hmcol <- rev(colorRampPalette(RColorBrewer::brewer.pal(9, "YlOrRd"))(255))
colnames(dists) <- rownames(expr.EGEOD18494.hif[expr.row])
diag(dists) <- NA 

ann_colors <- list(  
  time = RColorBrewer::brewer.pal(length(levels(data.EGEOD18494$time)), "Set2"),
  condition = c("#EF8A62", "#67A9CF")
)

ann_colors
names(ann_colors$time) <- levels(data.EGEOD18494$time)
names(ann_colors$condition) <- levels(data.EGEOD18494$condition)

p2 <- pheatmap(dists, #row = (hmcol), 
         #annotation_col = annotation_for_heatmap,
         #annotation_colors = ann_colors,
         legend = TRUE, 
         treeheight_row = 0,
         legend_breaks = c(min(dists, na.rm = TRUE), 
                           max(dists, na.rm = TRUE)), 
         legend_labels = (c("small dist", "large dist")),
         main = "Clustering of Gene Expression on Normoxia \n Euclidian Distance  (EGEOD18494)")

#plot_grid(p1, p2, labels = c('A', 'B'))

gridExtra::grid.arrange(grobs=list(p1$gtable, p2$gtable), 
                        nrow = 2 , labels=c('A', 'B'))

```

```{r}

data.EGEOD18494$time <- factor(data.EGEOD18494$time,  levels =  c("control", "4h" , "8h" , "12h 4h"))

row <- data.EGEOD18494$cell_line == "MDA-MB231 breast cancer"

annotation_for_heatmap <- droplevels(data.frame(time = data.EGEOD18494$time[row], condition = data.EGEOD18494$condition[row]))

row.names(annotation_for_heatmap) <- colnames(expr.EGEOD18494.hif)

#dists <- as.matrix(dist(expr.EGEOD18494.hif, method = "pearson")) 
dists <- cor(t(expr.EGEOD18494.hif), use = "pairwise.complete.obs", method = "spearman")


rownames(dists) <- rownames(expr.EGEOD18494.hif)

hmcol <- rev(colorRampPalette(RColorBrewer::brewer.pal(9, "YlOrRd"))(255))
colnames(dists) <- rownames(expr.EGEOD18494.hif)
diag(dists) <- NA 

ann_colors <- list(  
  time = RColorBrewer::brewer.pal(length(levels(data.EGEOD18494$time)), "Set2"),
  condition = c("#EF8A62", "#67A9CF")
)

ann_colors
names(ann_colors$time) <- levels(data.EGEOD18494$time)
names(ann_colors$condition) <- levels(data.EGEOD18494$condition)

pheatmap(dists, #row = (hmcol), 
         #annotation_col = annotation_for_heatmap,
         #annotation_colors = ann_colors,
         legend = TRUE, 
         treeheight_row = 0,
         legend_breaks = c(min(dists, na.rm = TRUE), 
                           max(dists, na.rm = TRUE)), 
         legend_labels = (c("-1", "1")),
         main = "Clustering of Gene Expression \n Spearman Correlation (EGEOD18494)")
```

