---
title: "BoolNet Inference ovarian cancer cell A2780 (GSE52695)"
output:
  pdf_document: default
  github_document: 
    df_print: paged
    html_preview: FALSE
    keep_html: FALSE
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_format = "all") })    
---

```{r error=TRUE, message=FALSE, warning=FALSE, include=FALSE, purl=FALSE, results='hide'}
## This chunk automatically generates a text .R version of this script when running within knitr.
# input  = knitr::current_input()  # filename of input document
# output = paste(tools::file_path_sans_ext(input), 'R', sep = '.')
# try(knitr::purl(input,output,documentation=2,quiet=T), silent = TRUE)
# # Avoid duplicate label error of knitr::purl
options(knitr.duplicate.label = 'allow')
# Code to browse the markdown file with renderized images.
knitr::opts_chunk$set(
  fig.path = "figs/GSE52695-"
)
```

RNA-seq analysis of Hypoxia and Normoxia cultured cancer cells (GSE52695)

Purpose: To study differential mRNA and precursor miRNA expression under hypoxia exposure in epithelial ovarian cancer cells. We treated ovarian cancer cell line A2780 under hypoxia condition at different time points. Normoxi acultured cells at corresponding time point were used as controls. Overall design 	A2780 cancer cells, cultured under hypoxia and normoxia. Time points tested: 6hr, 48hr, 6days.

https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE52695

Rupaimoole R, Wu SY, Pradeep S, Ivan C et al. Hypoxia-mediated downregulation of miRNA biogenesis promotes tumour progression. Nat Commun 2014 Oct 29;5:5202. PMID: 25351346

```{r message=FALSE, warning=FALSE}
packages_cran = c("igraph", "BoolNet", "BiocManager", "tidyverse", "fs", "effectsize")
# Install and load packages
package.check <- lapply(packages_cran, FUN = function(x) {
  if (!require(x, character.only = TRUE)) {
    install.packages(x, dependencies = TRUE)
    library(x, character.only = TRUE)
  }
})
packages_bioconductor = c("Biobase", "GEOquery", "vsn", "ArrayExpress", "illuminaHumanv3.db")
# Install and load packages
package.check <- lapply(packages_bioconductor, FUN = function(x) {
  if (!require(x, character.only = TRUE)) {
    BiocManager::install(x, dependencies = TRUE)
    library(x, character.only = TRUE)
  }
})
rm(package.check, packages_bioconductor, packages_cran)
```


<!-- ```{r} -->
<!-- download_dir <- fs::path(".data_tmp") -->
<!-- if (!dir_exists(download_dir)) { dir_create(download_dir) } -->

<!-- if (!dir_exists(".data_tmp/GSE52695_series_matrix.txt.gz")) { -->
<!--   GSE52695 <-getGEO("GSE52695", destdir = download_dir, GSEMatrix = T) -->
<!--   } else{ -->
<!--   GSE52695 <-getGEO(filename=".data_tmp/GSE52695_series_matrix.txt.gz") -->
<!--   } -->
<!-- GSE52695_supp <- getGEOSuppFiles("GSE52695", baseDir = download_dir) -->

<!-- expr.GSE52695 <- read.table(rownames(GSE52695_supp)[2], header = T, row.names = "ID") -->
<!-- prob.GSE52695 <- unique(rownames(expr.GSE52695)) -->

<!-- data.GSE52695 <- pData(GSE52695[[1]]) -->

<!-- data.GSE52695 <- data.GSE52695[data.GSE52695$description.1 %in% "RNA",] -->

<!-- expr.GSE52695 <- expr.GSE52695 %>% -->
<!--   rename_at(vars(data.GSE52695$description), ~data.GSE52695$geo_accession) -->

<!-- data.GSE52695 <- data.frame( -->
<!--                   codes = as.character(data.GSE52695$geo_accession), -->
<!--                   cell_line ="ovarian A2780", -->
<!--                   time = data.GSE52695$`time:ch1`, -->
<!--                   condition = data.GSE52695$`stress:ch1`) -->

<!-- data.GSE52695$rep <- rep(1:4, n= length(data.GSE52695$codes)) -->

<!-- data.GSE52695 <- data.GSE52695 %>% -->
<!--   mutate(time = recode(time, "6hr" = "6h", -->
<!--                              "48hr" = "2d", -->
<!--                              "6days" = "6d")) %>% -->
<!--   mutate(condition = recode(condition, "control" = "no", -->
<!--                              "Hypoxia" = "hy")) -->

<!-- rm(download_dir, GSE52695, GSE52695_supp, prob.GSE52695) -->

<!-- save.image("../data/data.GSE52695.Rdata") -->
<!-- ``` -->

# Load the pre-processed data

```{r}
load("../data/data.GSE52695.Rdata")
eset <- ExpressionSet(assayData = as.matrix(expr.GSE52695), 
                      probeNames = row.names(expr.GSE52695))
expr.GSE52695 <- exprs(justvsn(eset))
```

# Selecting the HIF Genes 
    
```{r}
# Selected genes from HIF Axis
hif.symbols <- c("TP53", "HIF1A", "EP300", "MDM2", "VHL")

# Select the probes and genes
expr.GSE52695.hif <- as.data.frame(expr.GSE52695) %>% 
  rownames_to_column('symbol') %>% 
  filter(symbol %in% hif.symbols) %>% 
  arrange(symbol)
```

# Example of Binarizing

```{r}
cols <- (data.GSE52695$rep == 1)
ov_1 <- 
expr.GSE52695.hif %>% 
  dplyr::select(c("symbol", data.GSE52695$codes[cols])) %>% arrange(symbol) %>% 
  arrange(symbol) %>% 
  rename_at(vars(data.GSE52695$codes[cols]),
            ~paste0(substr(data.GSE52695$condition[cols],1,2),".",
                    data.GSE52695$time[cols],".",
                    substr(data.GSE52695$cell_line[cols],1,2)))
knitr::kable(ov_1)

binarizeTimeSeries(ov_1[,-1], method="kmeans")$binarizedMeasurements  %>% 
  data.frame(.)  %>% 
  add_column(symbol = ov_1$symbol, .before=0) %>% 
  knitr::kable(.)
binarizeTimeSeries(ov_1[,-1], method="kmeans")$binarizedMeasurements  %>% 
  data.frame(.)  %>% 
  aggregate(., list(symbol = ov_1$symbol), mean) %>% 
  mutate_at(vars(-symbol), funs(ifelse(. >= 0.5, 1, 0))) %>% 
  rbind(., c("O2", 1,1,1)) %>% 
  knitr::kable(.)
```

```{r}
# Function to binarize according an consensus mean of probes, add the O2 state and rename columns 
binNetHy <- function(b){

  cols <- data.GSE52695$codes %in% names(b)

    binarizeTimeSeries(b[,-1], method="kmeans")$binarizedMeasurements  %>% 
    as.data.frame(.)  %>% 
    aggregate(., list(symbol = b$symbol), mean) %>%  # mean of binarized probes
    mutate_at(vars(-symbol), funs(ifelse(. >= 0.5, 1, 0))) %>%  # consensus with a bies to 1 (>= 0.5)
    rbind(., c("O2",0,0,0)) %>% 
      rename_at(vars(data.GSE52695$codes[cols] ),
              ~paste0(substr(data.GSE52695$condition[cols],1,2),".",
                      data.GSE52695$time[cols],".",
                      substr(data.GSE52695$cell_line[cols],1,2), ".",
                      data.GSE52695$rep[cols])) %>% 
    column_to_rownames("symbol")
}    
    
binNetNo <- function(b){    
    cols <- data.GSE52695$codes %in% names(b)

    binarizeTimeSeries(b[,-1], method="kmeans")$binarizedMeasurements  %>% 
    as.data.frame(.)  %>% 
    aggregate(., list(symbol = b$symbol), mean) %>%  # mean of binarized probes
    mutate_at(vars(-symbol), funs(ifelse(. >= 0.5, 1, 0))) %>%  # consensus with a bies to 1 (>= 0.5)
    rbind(., c("O2",1,1,1)) %>% 
      rename_at(vars(data.GSE52695$codes[cols] ),
              ~paste0(substr(data.GSE52695$condition[cols],1,2),".",
                      data.GSE52695$time[cols],".",
                      substr(data.GSE52695$cell_line[cols],1,2), ".",
                      data.GSE52695$rep[cols])) %>% 
    column_to_rownames("symbol")    
}

ov.1 <- 
expr.GSE52695.hif %>% 
  dplyr::select(c("symbol", data.GSE52695$codes[data.GSE52695$rep == 1]))  %>% 
  binNetNo(.) 

ov.2 <- 
expr.GSE52695.hif %>% 
  dplyr::select(c("symbol", data.GSE52695$codes[data.GSE52695$rep == 2]))  %>% 
  binNetNo(.) 

ov.3 <- 
expr.GSE52695.hif %>% 
  dplyr::select(c("symbol", data.GSE52695$codes[data.GSE52695$rep == 3]))  %>% 
  binNetHy(.) 

ov.4 <- 
expr.GSE52695.hif %>% 
  dplyr::select(c("symbol", data.GSE52695$codes[data.GSE52695$rep == 4]))  %>% 
  binNetHy(.) 

```

#  epithelial ovarian cancer cell line A2780

```{r}
# epithelial ovarian cancer cell line A2780 - 3 time-points, 4 replicates
ov.1.net <- reconstructNetwork(ov.1, method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)
ov.2.net <- reconstructNetwork(ov.2, method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)
ov.3.net <- reconstructNetwork(ov.3, method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)
ov.4.net <- reconstructNetwork(ov.4, method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)

ov.1.p <- plotNetworkWiring(ov.1.net, plotIt=F)
ov.2.p <- plotNetworkWiring(ov.2.net, plotIt=F)
ov.3.p <- plotNetworkWiring(ov.3.net, plotIt=F)
ov.4.p <- plotNetworkWiring(ov.4.net, plotIt=F)

# ov.all <- reconstructNetwork(list(ov.1.net,ov.2.net,ov.3.net,ov.4.net),method="bestfit",returnPBN=TRUE,readableFunctions=TRUE)
# ov.all.p <- plotNetworkWiring(ov.all, plotIt=F)
```

```{r}
# A2780 ovarian - 3 steps, replicate 1
print(ov.1.net)

# A2780 ovarian - 3 steps, replicate 2
print(ov.2.net)

# A2780 ovarian - 3 steps, replicate 3
print(ov.3.net)

# A2780 ovarian - 3 steps, replicate 4
print(ov.3.net)

# # A2780 ovarian - 3 steps, all replicates
# print(ov.all.p)
```


```{r}
# A2780 ovarian cancer - 4 time-points 
par(mfrow = c(1,4))
plot(ov.1.p, vertex.label.color="#440154ff", vertex.color="lightblue", vertex.frame.color="white", layout=layout_in_circle, edge.curved=.3,
     main="A2780 ovarian\n 3 steps, replicate 1")
plot(ov.2.p, vertex.label.color="#440154ff", vertex.color="lightblue", vertex.frame.color="white", layout=layout_in_circle, edge.curved=.3,
     main="A2780 ovarian\n 3 steps, replicate 2")
plot(ov.3.p, vertex.label.color="#440154ff", vertex.color="lightblue", vertex.frame.color="white", layout=layout_in_circle, edge.curved=.3,
     main="A2780 ovarian\n 3 steps, replicate 3")
plot(ov.4.p, vertex.label.color="#440154ff", vertex.color="lightblue", vertex.frame.color="white", layout=layout_in_circle, edge.curved=.3,
     main="A2780 ovarian\n 3 steps, replicate 4")

```

<!-- ```{r} -->
<!-- par(mfrow = c(1,1)) -->
<!-- plot(ov.all.p, vertex.label.color="#440154ff", vertex.color="lightblue", vertex.frame.color="white", layout=layout_in_circle, edge.curved=.3, -->
<!--      main="A2780 ovarian\n 3 steps, all replicates") -->
<!-- ``` -->

