---
title: "Attractors of IBNs"
author: "Patrick,Alexandre,Beatriz, Dhiego"
date: "16/12/2021"
output: pdf_document
---

---
title: "Attractors and GST"
author: "Patrick,  Alexandre, Diego e Beatriz"
date: "12/14/2021"
output: pdf_document
---

```{r}
library(BoolNet)
library(igraph)
```

```{r}
setwd("/storages/caico/home/alexandresarmento/R")
```
# GET ATTRACTOR AND GLOBAL STATE TRANSITIONS (GST) THRGOUGH IBN FROM ITERATIVE K-MEANS BINARIZATION

## ATTRACTOR IN SYNCRONOUS MODE (MDAMB231 IBN). 
```{r}
pathFile <- "/storages/caico/home/alexandresarmento/R/BNs/IBN_breast_KM2.bn"
booleanNetwork <- loadNetwork(pathFile)
attr <- getAttractors(booleanNetwork,type="synchronous")
plotAttractors(attr,
               drawLegend = FALSE,
               onColor = "#ffff00",
               offColor = "#0000ff",
               allInOnePlot=TRUE)
```
## MDAMB231 INITIAL CONDITIONS 
```{r}
load("/storages/caico/home/alexandresarmento/R/DataBases/breastkm2IC.Rdata")
```
## GST TOWARD TO ATTRACTORS (MDAMB231 IBN)
```{r}
basins1 <- initalConditions[[1]]
basins1[6] <- 0
basins1[10] <- 1
gst <- getPathToAttractor(booleanNetwork,basins1)
plotSequence(sequence=gst,
             drawLegend = FALSE,
             onColor = "#ffff00",
             offColor = "#0000ff")
```
## GST TOWARD TO ATTRACTORS (MDAMB231 IBN)
```{r}
basins2 <- initalConditions[[2]]
basins2[6] <- 0
basins2[10] <- 1
gst <- getPathToAttractor(booleanNetwork,basins2)
plotSequence(sequence=gst,
             drawLegend = FALSE,
             onColor = "#ffff00",
             offColor = "#0000ff")
```
## GST TOWARD TO ATTRACTORS (MDAMB231 IBN)
```{r}
basins3 <- initalConditions[[3]]
basins3[6] <- 0
basins3[10] <- 1
gst <- getPathToAttractor(booleanNetwork,basins3)
plotSequence(sequence=gst,
             drawLegend = FALSE,
             onColor = "#ffff00",
             offColor = "#0000ff")
```
## GST TOWARD TO ATTRACTORS (MDAMB231 IBN)
```{r}
basins4 <- initalConditions[[4]]
basins4[6] <- 0
basins4[10] <- 1
gst <- getPathToAttractor(booleanNetwork,basins4)
plotSequence(sequence=gst,
             drawLegend = FALSE,
             onColor = "#ffff00",
             offColor = "#0000ff")
```
## ATTRACTOR IN SYNCRONOUS MODE (HepG2 IBN)
```{r}
pathFile <- "/storages/caico/home/alexandresarmento/R/BNs/IBN_hepatoma_KM2.bn"
booleanNetwork <- loadNetwork(pathFile)
attrhepat <- getAttractors(booleanNetwork,type="synchronous")
plotAttractors(attrhepat,
               drawLegend = FALSE,
               onColor = "#ffff00",
               offColor = "#0000ff",
               allInOnePlot=TRUE)
```
## HepG2 INITIAL CONDITIONS 
```{r}
rm(initalConditions)
load("/storages/caico/home/alexandresarmento/R/DataBases/hepatomakm2IC.Rdata")
```
## GST TOWARD TO ATTRACTORS (HepG2 IBN)
```{r}
basins1 <- initalConditions[[1]]
basins1[7] <- 1 
basins1[11] <- 1 
gst <- getPathToAttractor(booleanNetwork,basins1)
plotSequence(sequence=gst,
             drawLegend = FALSE,
             onColor = "#ffff00",
             offColor = "#0000ff")

```
## GST TOWARD TO ATTRACTORS (HepG2 IBN)
```{r}
basins2 <- initalConditions[[2]]
basins2[7] <- 1 
basins2[11] <- 1 
gst <- getPathToAttractor(booleanNetwork,basins2)
plotSequence(sequence=gst,
             drawLegend = FALSE,
             onColor = "#ffff00",
             offColor = "#0000ff")
```
## GST TOWARD TO ATTRACTORS (HepG2 IBN)
```{r}
basins3 <- initalConditions[[3]]
basins3[7] <- 1 
basins3[11] <- 1 
gst <- getPathToAttractor(booleanNetwork,basins3)
plotSequence(sequence=gst,
             drawLegend = FALSE,
             onColor = "#ffff00",
             offColor = "#0000ff")
```
## GST TOWARD TO ATTRACTORS (HepG2 IBN)
```{r}
basins4 <- initalConditions[[4]]
basins4[7] <- 1 
basins4[11] <- 1 
gst <- getPathToAttractor(booleanNetwork,basins4)
plotSequence(sequence=gst,
             drawLegend = FALSE,
             onColor = "#ffff00",
             offColor = "#0000ff")
```
## ATTRACTOR IN SYNCRONOUS MODE (U87 IBN) 
```{r}
pathFile <- "/storages/caico/home/alexandresarmento/R/BNs/IBN_glioma_KM2.bn"
booleanNetwork <- loadNetwork(pathFile)
attr <- getAttractors(booleanNetwork,type="synchronous")
plotAttractors(attr,
               drawLegend = FALSE,
               onColor = "#ffff00",
               offColor = "#0000ff",
               allInOnePlot=TRUE)
```
## U87 INITIAL CONDITIONS 
```{r}
rm(initalConditions)
load("/storages/caico/home/alexandresarmento/R/DataBases/gliomakm2IC.Rdata")
```
## GST TOWARD TO ATTRACTORS (U87 IBN)
```{r}
basins1 <- initalConditions[[1]]
basins1[10] <- 1
basins1[12] <- 0
basins1[11] <- 1
basins1[9] <- 1 
gst <- getPathToAttractor(booleanNetwork,basins1)
plotSequence(sequence=gst,
             drawLegend = FALSE,
             onColor = "#ffff00",
             offColor = "#0000ff")
```
## GST TOWARD TO ATTRACTORS (U87 IBN)
```{r}
basins2 <- initalConditions[[2]]
basins2[10] <- 1
basins2[12] <- 0
basins2[11] <- 1
basins2[9] <- 1 
gst <- getPathToAttractor(booleanNetwork,basins2)
plotSequence(sequence=gst,
             drawLegend = FALSE,
             onColor = "#ffff00",
             offColor = "#0000ff")
```
## GST TOWARD TO ATTRACTORS (U87 IBN)
```{r}
basins3 <- initalConditions[[3]]
basins3[10] <- 1
basins3[12] <- 0
basins3[11] <- 1
basins3[9] <- 1 
gst <- getPathToAttractor(booleanNetwork,basins3)
plotSequence(sequence=gst,
             drawLegend = FALSE,
             onColor = "#ffff00",
             offColor = "#0000ff")

```
## GST TOWARD TO ATTRACTORS (U87 IBN)
```{r}
basins4 <- initalConditions[[4]]
basins4[10] <- 1
basins4[12] <- 0
basins4[11] <- 1
basins4[9] <- 1 
gst <- getPathToAttractor(booleanNetwork,basins4)
plotSequence(sequence=gst,
             drawLegend = FALSE,
             onColor = "#ffff00",
             offColor = "#0000ff")
```
# GET ATTRACTOR AND GLOBAL STATE TRANSITIONS (GST) THRGOUGH IBN FROM K-MEANS BINARIZATION

## ATTRACTOR IN SYNCRONOUS MODE (MDAMB231 IBN). 
```{r}
pathFile <- "/storages/caico/home/alexandresarmento/R/BNs/IBN_breast_KM1.bn"
booleanNetwork <- loadNetwork(pathFile)
attr <- getAttractors(booleanNetwork,type="synchronous")
plotAttractors(attr,
               drawLegend = FALSE,
               onColor = "#ffff00",
               offColor = "#0000ff",
               allInOnePlot=TRUE)
```
## MDAMB231 INITIAL CONDITIONS 
```{r}
rm(initalConditions)
load("/storages/caico/home/alexandresarmento/R/DataBases/breastkm1IC.Rdata")
```
## GST TOWARD TO ATTRACTORS (MDAMB231 IBN)
```{r}
basins1 <- initalConditions[[1]]
basins1[6] <- 0
basins1[10] <- 1
basins1[11] <- 1
gst <- getPathToAttractor(booleanNetwork,basins1)
plotSequence(sequence=gst,
             drawLegend = FALSE,
             onColor = "#ffff00",
             offColor = "#0000ff")
```
## GST TOWARD TO ATTRACTORS (MDAMB231 IBN)
```{r}
basins2 <- initalConditions[[2]]
basins2[6] <- 0
basins2[10] <- 1
basins2[11] <- 1
gst <- getPathToAttractor(booleanNetwork,basins2)
plotSequence(sequence=gst,
             drawLegend = FALSE,
             onColor = "#ffff00",
             offColor = "#0000ff")
```
## GST TOWARD TO ATTRACTORS (MDAMB231 IBN)
```{r}
basins3 <- initalConditions[[3]]
basins3[6] <- 0
basins3[10] <- 1
basins3[11] <- 1
gst <- getPathToAttractor(booleanNetwork,basins3)
plotSequence(sequence=gst,
             drawLegend = FALSE,
             onColor = "#ffff00",
             offColor = "#0000ff")
```
## GST TOWARD TO ATTRACTORS (MDAMB231 IBN)
```{r}
basins4 <- initalConditions[[4]]
basins4[6] <- 0
basins4[10] <- 1
basins4[11] <- 1
gst <- getPathToAttractor(booleanNetwork,basins4)
plotSequence(sequence=gst,
             drawLegend = FALSE,
             onColor = "#ffff00",
             offColor = "#0000ff")
```
## ATTRACTOR IN SYNCRONOUS MODE (HepG2 IBN)
```{r}
pathFile <- "/storages/caico/home/alexandresarmento/R/BNs/IBN_hepatoma_KM1.bn"
booleanNetwork <- loadNetwork(pathFile)
attr <- getAttractors(booleanNetwork,type="synchronous")
plotAttractors(attr,
               drawLegend = FALSE,
               onColor = "#ffff00",
               offColor = "#0000ff",
               allInOnePlot=TRUE)
```
## HepG2 INITIAL CONDITIONS 
```{r}
rm(initalConditions)
load("/storages/caico/home/alexandresarmento/R/DataBases/hepatomakm1IC.Rdata")
```
## GST TOWARD TO ATTRACTORS (HepG2 IBN)
```{r}
basins1 <- initalConditions[[1]]
basins1[6] <- 0
basins1[7] <- 1
basins1[11] <- 1
gst <- getPathToAttractor(booleanNetwork,basins1)
plotSequence(sequence=gst,
             drawLegend = FALSE,
             onColor = "#ffff00",
             offColor = "#0000ff")
```
## GST TOWARD TO ATTRACTORS (HepG2 IBN)
```{r}
basins2 <- initalConditions[[2]]
basins2[6] <- 0
basins2[7] <- 1
basins2[11] <- 1
gst <- getPathToAttractor(booleanNetwork,basins2)
plotSequence(sequence=gst,
             drawLegend = FALSE,
             onColor = "#ffff00",
             offColor = "#0000ff")
```
## GST TOWARD TO ATTRACTORS (HepG2 IBN)
```{r}
basins3 <- initalConditions[[3]]
basins3[6] <- 0
basins3[7] <- 1
basins3[11] <- 1
gst <- getPathToAttractor(booleanNetwork,basins3)
plotSequence(sequence=gst,
             drawLegend = FALSE,
             onColor = "#ffff00",
             offColor = "#0000ff")
```
## GST TOWARD TO ATTRACTORS (HepG2 IBN)
```{r}
basins4 <- initalConditions[[4]]
basins4[6] <- 0
basins4[7] <- 1
basins4[11] <- 1
gst <- getPathToAttractor(booleanNetwork,basins4)
plotSequence(sequence=gst,
             drawLegend = FALSE,
             onColor = "#ffff00",
             offColor = "#0000ff")
```
## ATTRACTOR IN SYNCRONOUS MODE (U87 IBN) 
```{r}
pathFile <- "/storages/caico/home/alexandresarmento/R/BNs/IBN_glioma_KM1.bn"
booleanNetwork <- loadNetwork(pathFile)
attr <- getAttractors(booleanNetwork,type="synchronous")
plotAttractors(attr,
               drawLegend = FALSE,
               onColor = "#ffff00",
               offColor = "#0000ff",
               allInOnePlot=TRUE)
```
## U87 INITIAL CONDITIONS 
```{r}
rm(initalConditions)
load("/storages/caico/home/alexandresarmento/R/DataBases/gliomakm1IC.Rdata")
```
## GST TOWARD TO ATTRACTORS (U87 IBN)
```{r}
basins1 <- initalConditions[[1]]
basins1[2] <- 0
basins1[6] <- 0
basins1[9] <- 1
basins1[10] <- 1
basins1[11] <- 1
basins1[12] <- 0
gst <- getPathToAttractor(booleanNetwork,basins1)
plotSequence(sequence=gst,
             drawLegend = FALSE,
               onColor = "#ffff00",
               offColor = "#0000ff")
```
## GST TOWARD TO ATTRACTORS (U87 IBN)
```{r}
basins2 <- initalConditions[[2]]
basins2[2] <- 0
basins2[6] <- 0
basins2[9] <- 1
basins2[10] <- 1
basins2[11] <- 1
basins2[12] <- 0
gst <- getPathToAttractor(booleanNetwork,basins2)
plotSequence(sequence=gst,
             drawLegend = FALSE,
               onColor = "#ffff00",
               offColor = "#0000ff")
```
## GST TOWARD TO ATTRACTORS (U87 IBN)
```{r}
basins3 <- initalConditions[[3]]
basins3[2] <- 0
basins3[6] <- 0
basins3[9] <- 1
basins3[10] <- 1
basins3[11] <- 1
basins3[12] <- 0
gst <- getPathToAttractor(booleanNetwork,basins3)
plotSequence(sequence=gst,
             drawLegend = FALSE,
               onColor = "#ffff00",
               offColor = "#0000ff")
```
## GST TOWARD TO ATTRACTORS (U87 IBN)
```{r}
basins4 <- initalConditions[[4]]
basins4[2] <- 0
basins4[6] <- 0
basins4[9] <- 1
basins4[10] <- 1
basins4[11] <- 1
basins4[12] <- 0
gst <- getPathToAttractor(booleanNetwork,basins4)
plotSequence(sequence=gst,
             drawLegend = FALSE,
               onColor = "#ffff00",
               offColor = "#0000ff")
```

# GET ATTRACTOR AND GLOBAL STATE TRANSITIONS (GST) THRGOUGH IBN FROM BASCA_TAU =0.2 BINARIZATION

## ATTRACTOR IN SYNCRONOUS MODE (MDAMB231 IBN). 
```{r}
pathFile <- "/storages/caico/home/alexandresarmento/R/BNs/IBN_breast_basca.bn"
booleanNetwork <- loadNetwork(pathFile)
attr <- getAttractors(booleanNetwork,type="synchronous")
plotAttractors(attr,
               drawLegend = FALSE,
               onColor = "#ffff00",
               offColor = "#0000ff",
               allInOnePlot=TRUE)
```
## MDAMB231 INITIAL CONDITIONS 
```{r}
rm(initalConditions)
load("/storages/caico/home/alexandresarmento/R/DataBases/breastbascaIC.Rdata")
```
## GST TOWARD TO ATTRACTORS (MDAMB231 IBN)
```{r}
basins1 <- initalConditions[[1]]
basins1[6] <- 0
basins1[10] <- 1
basins1[11] <- 1
gst <- getPathToAttractor(booleanNetwork,basins1)
plotSequence(sequence=gst,
             drawLegend = FALSE,
             onColor = "#ffff00",
             offColor = "#0000ff")
```
## GST TOWARD TO ATTRACTORS (MDAMB231 IBN)
```{r}
basins2 <- initalConditions[[2]]
basins2[6] <- 0
basins2[10] <- 1
basins2[11] <- 1
gst <- getPathToAttractor(booleanNetwork,basins2)
plotSequence(sequence=gst,
             drawLegend = FALSE,
             onColor = "#ffff00",
             offColor = "#0000ff")
```
## GST TOWARD TO ATTRACTORS (MDAMB231 IBN)
```{r}
basins3 <- initalConditions[[3]]
basins3[6] <- 0
basins3[10] <- 1
basins3[11] <- 1
gst <- getPathToAttractor(booleanNetwork,basins3)
plotSequence(sequence=gst,
             drawLegend = FALSE,
             onColor = "#ffff00",
             offColor = "#0000ff")
```
## GST TOWARD TO ATTRACTORS (MDAMB231 IBN)
```{r}
basins4 <- initalConditions[[4]]
basins4[6] <- 0
basins4[10] <- 1
basins4[11] <- 1
gst <- getPathToAttractor(booleanNetwork,basins4)
plotSequence(sequence=gst,
             drawLegend = FALSE,
             onColor = "#ffff00",
             offColor = "#0000ff")
```
## ATTRACTOR IN SYNCRONOUS MODE (HepG2 IBN)
```{r}
pathFile <- "/storages/caico/home/alexandresarmento/R/BNs/IBN_hepatoma_basca.bn"
booleanNetwork <- loadNetwork(pathFile)
attr <- getAttractors(booleanNetwork,type="synchronous")
plotAttractors(attr,
               drawLegend = FALSE,
               onColor = "#ffff00",
               offColor = "#0000ff",
               allInOnePlot=TRUE)
```
## HepG2 INITIAL CONDITIONS 
```{r}
rm(initalConditions)
load("/storages/caico/home/alexandresarmento/R/DataBases/hepatomabascaIC.Rdata")
```
## GST TOWARD TO ATTRACTORS (HepG2 IBN)
```{r}
basins1 <- initalConditions[[1]]
basins1[6] <- 0
basins1[7] <- 1
basins1[11] <- 1
gst <- getPathToAttractor(booleanNetwork,basins1)
plotSequence(sequence=gst,
             drawLegend = FALSE,
               onColor = "#ffff00",
               offColor = "#0000ff")
```
## GST TOWARD TO ATTRACTORS (HepG2 IBN)
```{r}
basins2 <- initalConditions[[2]]
basins2[6] <- 0
basins2[7] <- 1
basins2[11] <- 1
gst <- getPathToAttractor(booleanNetwork,basins2)
plotSequence(sequence=gst,
             drawLegend = FALSE,
               onColor = "#ffff00",
               offColor = "#0000ff")
```
## GST TOWARD TO ATTRACTORS (HepG2 IBN)
```{r}
basins3 <- initalConditions[[3]]
basins3[6] <- 0
basins3[7] <- 1
basins3[11] <- 1
gst <- getPathToAttractor(booleanNetwork,basins3)
plotSequence(sequence=gst,
             drawLegend = FALSE,
               onColor = "#ffff00",
               offColor = "#0000ff")
```
## GST TOWARD TO ATTRACTORS (HepG2 IBN)
```{r}
basins4 <- initalConditions[[4]]
basins4[6] <- 0
basins4[7] <- 1
basins4[11] <- 1
gst <- getPathToAttractor(booleanNetwork,basins4)
plotSequence(sequence=gst,
             drawLegend = FALSE,
               onColor = "#ffff00",
               offColor = "#0000ff")
```
## ATTRACTOR IN SYNCRONOUS MODE (U87 IBN) 
```{r}
pathFile <- "/storages/caico/home/alexandresarmento/R/BNs/IBN_glioma_basca.bn"
booleanNetwork <- loadNetwork(pathFile)
attr <- getAttractors(booleanNetwork,type="synchronous")
plotAttractors(attr,
               drawLegend = FALSE,
               onColor = "#ffff00",
               offColor = "#0000ff",
               allInOnePlot=TRUE)
```
## U87 INITIAL CONDITIONS 
```{r}
rm(initalConditions)
load("/storages/caico/home/alexandresarmento/R/DataBases/gliomabascaIC.Rdata")
```
## GST TOWARD TO ATTRACTORS (U87 IBN)
```{r}
basins1 <- initalConditions[[1]]
basins1[2] <- 0
basins1[6] <- 0
basins1[9] <- 1
basins1[10] <- 1
basins1[11] <- 1
basins1[12] <- 0
gst <- getPathToAttractor(booleanNetwork,basins1)
plotSequence(sequence=gst,
             drawLegend = FALSE,
               onColor = "#ffff00",
               offColor = "#0000ff")
```
## GST TOWARD TO ATTRACTORS (U87 IBN)
```{r}
basins2 <- initalConditions[[2]]
basins2[2] <- 0
basins2[6] <- 0
basins2[9] <- 1
basins2[10] <- 1
basins2[11] <- 1
basins2[12] <- 0
gst <- getPathToAttractor(booleanNetwork,basins2)
plotSequence(sequence=gst,
             drawLegend = FALSE,
               onColor = "#ffff00",
               offColor = "#0000ff")
```
## GST TOWARD TO ATTRACTORS (U87 IBN)
```{r}
basins3 <- initalConditions[[3]]
basins3[2] <- 0
basins3[6] <- 0
basins3[9] <- 1
basins3[10] <- 1
basins3[11] <- 1
basins3[12] <- 0
gst <- getPathToAttractor(booleanNetwork,basins3)
plotSequence(sequence=gst,
             drawLegend = FALSE,
               onColor = "#ffff00",
               offColor = "#0000ff")
```
## GST TOWARD TO ATTRACTORS (U87 IBN)
```{r}
basins4 <- initalConditions[[4]]
basins4[2] <- 0
basins4[6] <- 0
basins4[9] <- 1
basins4[10] <- 1
basins4[11] <- 1
basins4[12] <- 0
gst <- getPathToAttractor(booleanNetwork,basins4)
plotSequence(sequence=gst,
             drawLegend = FALSE,
               onColor = "#ffff00",
               offColor = "#0000ff")
```


# GET ATTRACTOR AND GLOBAL STATE TRANSITIONS (GST) THRGOUGH IBN FROM BASCB_TAU =0.2 BINARIZATION

## ATTRACTOR IN SYNCRONOUS MODE (MDAMB231 IBN). 
```{r}
pathFile <- "/storages/caico/home/alexandresarmento/R/BNs/IBN_breast_bascb.bn"
booleanNetwork <- loadNetwork(pathFile)
attr <- getAttractors(booleanNetwork,type="synchronous")
plotAttractors(attr,
               drawLegend = FALSE,
               onColor = "#ffff00",
               offColor = "#0000ff",
               allInOnePlot=TRUE)
```
## MDAMB231 INITIAL CONDITIONS 
```{r}
rm(initalConditions)
load("/storages/caico/home/alexandresarmento/R/DataBases/breastbascbIC.Rdata")
```
## GST TOWARD TO ATTRACTORS (MDAMB231 IBN)
```{r}
basins1 <- initalConditions[[1]]
basins1[6] <- 0
basins1[10] <- 1
basins1[11] <- 1
gst <- getPathToAttractor(booleanNetwork,basins1)
plotSequence(sequence=gst,
             drawLegend = FALSE,
             onColor = "#ffff00",
             offColor = "#0000ff")
```
## GST TOWARD TO ATTRACTORS (MDAMB231 IBN)
```{r}
basins2 <- initalConditions[[2]]
basins2[6] <- 0
basins2[10] <- 1
basins2[11] <- 1
gst <- getPathToAttractor(booleanNetwork,basins2)
plotSequence(sequence=gst,
             drawLegend = FALSE,
             onColor = "#ffff00",
             offColor = "#0000ff")
```
## GST TOWARD TO ATTRACTORS (MDAMB231 IBN)
```{r}
basins3 <- initalConditions[[3]]
basins3[6] <- 0
basins3[10] <- 1
basins3[11] <- 1
gst <- getPathToAttractor(booleanNetwork,basins3)
plotSequence(sequence=gst,
             drawLegend = FALSE,
             onColor = "#ffff00",
             offColor = "#0000ff")
```
## GST TOWARD TO ATTRACTORS (MDAMB231 IBN)
```{r}
basins4 <- initalConditions[[4]]
basins4[6] <- 0
basins4[10] <- 1
basins4[11] <- 1
gst <- getPathToAttractor(booleanNetwork,basins4)
plotSequence(sequence=gst,
             drawLegend = FALSE,
             onColor = "#ffff00",
             offColor = "#0000ff")
```
## ATTRACTOR IN SYNCRONOUS MODE (HepG2 IBN)
```{r}
pathFile <- "/storages/caico/home/alexandresarmento/R/BNs/IBN_hepatoma_bascb.bn"
booleanNetwork <- loadNetwork(pathFile)
attr <- getAttractors(booleanNetwork,type="synchronous")
plotAttractors(attr,
               drawLegend = FALSE,
               onColor = "#ffff00",
               offColor = "#0000ff",
               allInOnePlot=TRUE)
```
## HepG2 INITIAL CONDITIONS 
```{r}
rm(initalConditions)
load("/storages/caico/home/alexandresarmento/R/DataBases/hepatomabascbIC.Rdata")
```
## GST TOWARD TO ATTRACTORS (HepG2 IBN)
```{r}
basins1 <- initalConditions[[1]]
basins1[6] <- 0
basins1[7] <- 1
basins1[11] <- 1
gst <- getPathToAttractor(booleanNetwork,basins1)
plotSequence(sequence=gst,
             drawLegend = FALSE,
               onColor = "#ffff00",
               offColor = "#0000ff")
```
## GST TOWARD TO ATTRACTORS (HepG2 IBN)
```{r}
basins2 <- initalConditions[[2]]
basins2[6] <- 0
basins2[7] <- 1
basins2[11] <- 1
gst <- getPathToAttractor(booleanNetwork,basins2)
plotSequence(sequence=gst,
             drawLegend = FALSE,
               onColor = "#ffff00",
               offColor = "#0000ff")
```
## GST TOWARD TO ATTRACTORS (HepG2 IBN)
```{r}
basins3 <- initalConditions[[3]]
basins3[6] <- 0
basins3[7] <- 1
basins3[11] <- 1
gst <- getPathToAttractor(booleanNetwork,basins3)
plotSequence(sequence=gst,
             drawLegend = FALSE,
               onColor = "#ffff00",
               offColor = "#0000ff")
```
## GST TOWARD TO ATTRACTORS (HepG2 IBN)
```{r}
basins4 <- initalConditions[[4]]
basins4[6] <- 0
basins4[7] <- 1
basins4[11] <- 1
gst <- getPathToAttractor(booleanNetwork,basins4)
plotSequence(sequence=gst,
             drawLegend = FALSE,
               onColor = "#ffff00",
               offColor = "#0000ff")
```
## ATTRACTOR IN SYNCRONOUS MODE (U87 IBN) 
```{r}
pathFile <- "/storages/caico/home/alexandresarmento/R/BNs/IBN_glioma_bascb.bn"
booleanNetwork <- loadNetwork(pathFile)
attr <- getAttractors(booleanNetwork,type="synchronous")
plotAttractors(attr,
               drawLegend = FALSE,
               onColor = "#ffff00",
               offColor = "#0000ff",
               allInOnePlot=TRUE)
```
## U87 INITIAL CONDITIONS 
```{r}
rm(initalConditions)
load("/storages/caico/home/alexandresarmento/R/DataBases/gliomabascbIC.Rdata")
```
## GST TOWARD TO ATTRACTORS (U87 IBN)
```{r}
basins1 <- initalConditions[[1]]
basins1[2] <- 0
basins1[6] <- 0
basins1[9] <- 1
basins1[10] <- 1
basins1[11] <- 1
basins1[12] <- 0
gst <- getPathToAttractor(booleanNetwork,basins1)
plotSequence(sequence=gst,
             drawLegend = FALSE,
               onColor = "#ffff00",
               offColor = "#0000ff")
```
## GST TOWARD TO ATTRACTORS (U87 IBN)
```{r}
basins2 <- initalConditions[[2]]
basins2[2] <- 0
basins2[6] <- 0
basins2[9] <- 1
basins2[10] <- 1
basins2[11] <- 1
basins2[12] <- 0
gst <- getPathToAttractor(booleanNetwork,basins2)
plotSequence(sequence=gst,
             drawLegend = FALSE,
               onColor = "#ffff00",
               offColor = "#0000ff")
```
## GST TOWARD TO ATTRACTORS (U87 IBN)
```{r}
basins3 <- initalConditions[[3]]
basins3[2] <- 0
basins3[6] <- 0
basins3[9] <- 1
basins3[10] <- 1
basins3[11] <- 1
basins3[12] <- 0
gst <- getPathToAttractor(booleanNetwork,basins3)
plotSequence(sequence=gst,
             drawLegend = FALSE,
               onColor = "#ffff00",
               offColor = "#0000ff")
```
## GST TOWARD TO ATTRACTORS (U87 IBN)
```{r}
basins4 <- initalConditions[[4]]
basins4[2] <- 0
basins4[6] <- 0
basins4[9] <- 1
basins4[10] <- 1
basins4[11] <- 1
basins4[12] <- 0
gst <- getPathToAttractor(booleanNetwork,basins4)
plotSequence(sequence=gst,
             drawLegend = FALSE,
               onColor = "#ffff00",
               offColor = "#0000ff")
```